### 2015-04-08 draft.tex
..

\documentclass[12pt]{article}
\usepackage{times}
\usepackage{latexsym}
\usepackage[bottom=20mm, top=20mm, left=20mm, right=20mm]{geometry}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{bussproofs}
\usepackage{comment}
\usepackage[numbers]{natbib}
\usepackage{graphicx} % Graphics support
\usepackage{caption} % Improved captions
\usepackage{subfig} % Inclusion of subfigures
%\usepackage{subfigure}

% adjust space between letters
\usepackage{microtype}

% adjust space between lines
\linespread{1.05}

% for header and footer
\usepackage{fancyhdr}

% change the font size of section
%\usepackage{sectsty}
%\sectionfont{\fontsize{12}{15}\selectfont}
%\subsectionfont{\fontsize{11}{14}\selectfont}

\usepackage{amsmath, amssymb}

\pagestyle{fancy}

\lhead{Department of Computer Science\\
Princeton University}
\chead{}
% \rhead{\includegraphics[height=15mm]{princeton_logo.jpg}}
\lfoot{}
\cfoot{}
\rfoot{}

% remove page numbers
%\thispagestyle{empty}

% no indent at the start of paragraphs
\parindent=0em

% for increasing space between paragraphs
\newcommand{\emptyline}{\vspace{5pt}}

%-----------------------------------------------------------

\begin{document}
\quad \\
\quad \\
\quad \\
\centerline{ \large{Towards Range-Based Network Synthesis}}
\vspace{1pt}\\
\centerline{Ryan Beckett and Olivier Savary B\'{e}langer} \\
\\
\centerline{April 15, 2015} \\

\section*{Introduction}
% Including an example policy

\section*{Algorithic Policies Revisited}
% Explains what algorithmic policies are, how maple was handling them and the difference with our model

\section*{API and Decision Trees}
   \subsection*{Range-based Branching}
   % Why range instead of actual value, pre-generation of rules using symbolic pre-execution

   \subsection*{Built-in Relation with Automatic Invalidation}
   % Relation with add and remove tuple of fields from a relation (empty at the begining), Inhabitant-based Branching

   % Computing which packets need to be sent to the controller (i.e. only the ones reaching an ``add'' or a ``remove'' node.

\section*{Implementation}
%that's the hope...

   \subsection*{Optimizations}
   % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
   % 2) Prune dead subtree on empty ranges and add/remove which do not affect any inhabitant-based branching.

\section*{Evaluation}
% 1) optimality w.r.t. packets seen by the controler?
% 2) more examples
% 3) simulated run

\section*{Related Work}
% deeper comparison with Maple

%???

\section*{Future Work and Conclusion}

\bibliographystyle{plainnat}
\bibliography{bibi}

\end{document}

### 2015-04-08 proposal.tex
..

\documentclass[12pt]{article}
\usepackage{times}
\usepackage{latexsym}
\usepackage[bottom=20mm, top=20mm, left=20mm, right=20mm]{geometry}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{bussproofs}
\usepackage{comment}
\usepackage[numbers]{natbib}
\usepackage{graphicx} % Graphics support
\usepackage{caption} % Improved captions
\usepackage{subfig} % Inclusion of subfigures
%\usepackage{subfigure}

% adjust space between letters
\usepackage{microtype}

% adjust space between lines
\linespread{1.05}

% for header and footer
\usepackage{fancyhdr}

% change the font size of section
%\usepackage{sectsty}
%\sectionfont{\fontsize{12}{15}\selectfont}
%\subsectionfont{\fontsize{11}{14}\selectfont}

\usepackage{amsmath, amssymb}

\pagestyle{fancy}

\lhead{Department of Computer Science\\
Princeton University}
\chead{}
% \rhead{\includegraphics[height=15mm]{princeton_logo.jpg}}
\lfoot{}
\cfoot{}
\rfoot{}

% remove page numbers
%\thispagestyle{empty}

% no indent at the start of paragraphs
\parindent=0em

% for increasing space between paragraphs
\newcommand{\emptyline}{\vspace{5pt}}

%-----------------------------------------------------------

\begin{document}
\quad \\
\quad \\
\quad \\
\centerline{ \large{COS 561 Proposal}}
\vspace{1pt}\\
\centerline{Ryan Beckett and Olivier Savary Belanger} \\
\\
\centerline{March 24, 2015} \\

\section*{Introduction}

Software-Defined Networking (SDN) has enabled centralized network programmability by providing a simple, open interface to the hardware. Maple \cite{Maple} simplifies the task of writing centralized network programs by allowing operators to write a forwarding decision program which conceptually runs on every packet in the network. To achieve controller scalability, Maple records previous decisions on packets and installs forwarding rules to handle similar future traffic directly in the data plane. However, Maple is reactive and must witness how the controller handles a packet before installing forwarding rules, forcing all traffic to initially go through the controller. Also, Maple lacks an elegant mechanism for handling stateful programs, relying instead on low-level, user-supplied trace invalidation schemes. We plan to address both of these limitations by redesigning the Maple API to proactively generate and generalize rules from a controller program to avoid initial packet latency, and to automate invalidation for stateful controller programs by providing a built in table data structure.

\section*{Generalization and Proactive Generation of Trace Tree}

%	+ Generalization made possible by range-based queries rather than returning the field's value (get)
\paragraph{A New API for More Efficient Trace Tree} Maple uses a trace tree to memoize the results of applying the algorithmic policy to a class of packets. Nodes in the tree corresponds to calls to the Maple API, and each child corresponds to a return value of the API call. In Maple, the trace tree starts empty and expands when the controller processes a packet whose trace is not present in the tree. 
However, we observe that it is infrequent for a controller to apply different policies to each different values of a field. Rather, the same policy is used for a range of values (e.g., if a packet has a certain IP prefix). A call to the Maple API would often be followed by an if-statement with a boolean predicate which depends on the returned value. In that case, the tree could be compressed by internalizing the boolean predicate in the API. Such an API would consists of a function from ranges to Boolean for each fields in a packet. As with the current Maple API, all decisions involving packets field values are captured by the trace tree, making this API fully language-agnostic. We believe the two APIs to be of equivalent strength.
%(TODONote: as powerful, translation between "i <- get(dst); case i -> f_i" and "while(i++) if r(i) => case(i);break with function case(i) case i -> f_i" equivalent to case-semantics)

% + Proactive generation that explores the complement of the ranges encountered (similar to directed random testing). As all decision involving the field's values are captured by the range-based queries, this remains language agnostic.
\paragraph{Proactive Generation of Trace Tree} This revised API allows exploring the trace tree from a single packet (possibly randomly generated). At each API call encountered, we would be generating a symbolic packet representing a value in the complement of the field range encountered. This could be done ahead of time, eliminating the start-up phase of Maple where most packets are missed with respect to the trace tree.

%	+ Model makes easy to express constraints over the domain of possible packets (e.g. certain networks may only deal with hosts in a certain IP range)
\paragraph{Constrained Topology and Packet Domain} The API's model allows for easy expression of constraints over the domain of possible packets, representing them as a series of nodes rooting the trace tree, each constraining the possible range of a field, with the false portion of the tree marked as unreachable.

\section*{Stateful controller program}
%	+ Relation data structure to express controller state that persists across controller executions. (e.g., can express MAC learning table, stateful firewall allowed hosts).

\paragraph{State as Mutable Relations}
Maple supports stateful controller programs through algorithm policies taking in an environment in addition to a packet and returning a forwarding decision and a modified environment. However, by allowing environments to be arbitrary data structures, the burden is on the programmer to invalidate the appropriate traces. Invalidating too many traces would be inefficient, while invalidating too few would be unsound. We would like to include in our API a relational data structure that programmers can use to capture the current state of the controller program. These principled environments would benefit from a built-in invalidation mechanism. We believe that this simple abstraction is general enough to represent most examples of environment used in real-world controller programs, for example a MAC-learning table, a firewall ACL or a mutable topology.

\paragraph{State in the Trace Tree}
Because all state modifications would be performed through our API, we can capture these in the trace tree.  
By analyzing the modifications of relation data structures through the API (adding and removing elements), we can further inform proactive generation. For example, if only certain ranges of field values can ever be added to a relation, then we can proactively generate rules to handle the ranges of values that can never appear in the relation. Also, we can discover through API calls that some relations, such as stateful firewall ACL, are monotonic.

%\paragraph{Automatic Invalidation of Traces}

\section*{Estimated Timeline}
\begin{itemize}
\item By mid-April, we want to have:
  \begin{itemize}
  \item completed the design of the API
  \item and devised a sound invalidation mechanism for traces affected by mutable relations
  \end{itemize}

\item By the end of April, we want to have:
  \begin{itemize}
  \item a working implementation
  \item preliminary evaluation including examples of stateful controller programs and percentage of packets forwarded to the controller
  \end{itemize}
\end{itemize}

\bibliographystyle{plainnat}
\bibliography{bibi}

\end{document}

### 2015-04-12 draft.tex
..

@@
 \usepackage{enumitem}
+\usepackage{listings}
 \usepackage{bussproofs}
@@
 \quad \\
-\centerline{ \large{Towards Range-Based Network Synthesis}}
+\centerline{ \large{Red Oak: Something Something Algorithmic Policy}}
 \vspace{1pt}\\
@@
 \section*{Introduction}
+% possibly in introduction instead?
+Maple \cite{Maple} proposes a SDN controller model where the behavior of the controller is defined by a function from packets to forwarding path, the algorithmic policy.
+The authors then show how to optimize the controller so that packets are handled without having to run the function on them.
+Maple does so by providing a packet API capturing each access to packet fields, and memoizing the results of the function for subsequent packets sharing the accessed fields' values.
+We identify a number of shortcoming to this model:
+\begin{enumerate}
+\item % packet value granularity too fine
+  \item % environmentgranularity too fine, no built-in support for dynamic controllers
+  \item % not proactive
+\end{enumerate}
+
+We present a new iteration of the model where
+\begin{enumerate}
+  \item calls to our API does not expose the values of a packet's field, answering instead boolean queries about the values being in a certain range
+  \item dynamic controllers are supported through a relation data-structure which is updated through an extended API
+  \item the trace tree can be generated from executing the policy on a single (potentially symbolic) packet.
+\end{enumerate}
+
 % Including an example policy
+
+\begin{figure}
+\begin{lstlisting}
+f(pkt):
+  if RO.InRange(pkt, src, 0, 10) then
+     RO.add(pkt, (dst, src), conn);
+     fw(gateway)
+  else if (RO.InSet(pkt, (src,dst), conn) then
+         fw(dst)   
+       else
+         drop
+       end
+  end                
+  \end{lstlisting}
+
+\caption{A Stateful Firewall}
+\end{figure}
 
@@
 
+   \begin{figure}
+\begin{lstlisting}
+Module RO
+
+Type pkt (* contains a range for each field *)
+
+Type field (* enum each field *)
+
+function InRange: pkt * field * value * value -> bool
+
+\end{lstlisting}
+
+\caption{Signature of the packet portion of the API}
+\end{figure}
+
+   % Can represent constraints on the ranges admissible given a certain topology as a series of ``true'' branches at the root of the tree.
+
    \subsection*{Built-in Relation with Automatic Invalidation}
    % Relation with add and remove tuple of fields from a relation (empty at the begining), Inhabitant-based Branching
+
+   \begin{figure}
+     \begin{lstlisting}
+
+function InRange: pkt * field * value * value -> bool
+
+\end{lstlisting}
+
+\caption{Signature of the relation portion of the API}
+\end{figure}
 

### 2015-04-15 draft.tex
..

@@
 
-We present a new iteration of the model where
+To address these, we modify the model in the following way:
 \begin{enumerate}
@@
 
-% Including an example policy
+% Including an example policy, todo change 0 to 10 to actual IP
+Figure~\ref{fig:firewall-eg} displays a policy for a stateful firewall using our API.
+Sources from 0 to 10 are taken to be internal to the network.
+If a packet is sent from an internal host to an external host, the pair of internal and external host is added to a list of trusted connection.
+If a packet is received from an external host, we lookup the table to see if the connection is trusted, in which case we forward the packet to its destination, otherwise we drop it.
 
@@
   if RO.InRange(pkt, src, 0, 10) then
-     RO.add(pkt, (dst, src), conn);
-     fw(gateway)
-  else if (RO.InSet(pkt, (src,dst), conn) then
-         fw(dst)   
+     if RO.InRange(pkt, dst, 0, 10) then
+        RO.deliver()
      else 
-         drop
+        RO.add(pkt, (dst, src), trusted);
+        RO.deliver()
+  else if (RO.InSet(pkt, (src,dst), trusted) then
+         RO.deliver()
+       else
+         RO.drop
        end
@@
    % Can represent constraints on the ranges admissible given a certain topology as a series of ``true'' branches at the root of the tree.
+
+   \subsection*{Forwarding Decision}
+   % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to dst)
 

### 2015-04-17 draft.tex
..

@@
 \section*{Introduction}
-% possibly in introduction instead?
+
 Maple \cite{Maple} proposes a SDN controller model where the behavior of the controller is defined by a function from packets to forwarding path, the algorithmic policy.
@@
 \begin{enumerate}
-\item % packet value granularity too fine
-  \item % environmentgranularity too fine, no built-in support for dynamic controllers
-  \item % not proactive
+\item When a field value is observe, the forwarding decision only works on packets with this exact value, even if the forwarding decision is made on a much coarser range of value.Similarly, after observing some value in the environment, the forwarding decision is specialize for that particular value
+
+\item It no built-in environment datastructure with automatic invalidation of decision trace when the environment is modified, which is needed to implement dynamic controllers.
+
+  \item Populating the decision tree requires seeing a high number of packets, potentially slowing down the network at start-up. 
 \end{enumerate}
@@
 \end{enumerate}
+
+\section*{Algorithic Policies Revisited}
+% Explains what algorithmic policies are, how maple was handling them and the difference with our model 
+Algorithmic policies were introduced by Maple as a model for SDN controllers.
+An algorithmic policy consist of a function from packet to forwarding decision.
+Under this model, the network forwards packets as if this function was executed on each packet when entering the network.
 
@@
 f(pkt):
-  if RO.InRange(pkt, src, 0, 10) then
-     if RO.InRange(pkt, dst, 0, 10) then
+  if RO.in_range(pkt, src, 0, 10) then
+     if RO.in_range(pkt, dst, 0, 10) then
         RO.deliver()
@@
         RO.deliver()
-  else if (RO.InSet(pkt, (src,dst), trusted) then
+  else if (RO.in_set(pkt, (src,dst), trusted) then
          RO.deliver()
@@
 \end{figure}
-
-\section*{Algorithic Policies Revisited}
-% Explains what algorithmic policies are, how maple was handling them and the difference with our model
 

### 2015-04-26 draft.tex
..

@@
 \documentclass[12pt]{article}
+\usepackage[latin1]{inputenc}
+
 \usepackage{times}
@@
 \usepackage{subfig} % Inclusion of subfigures
+
+\usepackage{tikz}
+\usepackage{tikz-qtree}
+
+% Set the overall layout of the tree
+%\tikzstyle{level 1}=[level distance=3.5cm, sibling distance=3.5cm]
+%\tikzstyle{level 2}=[level distance=3.5cm, sibling distance=2cm]
+
+% Define styles for bags and leafs
+%\tikzstyle{leaf} = [draw=none,fill=none, text width=4em, text centered]
+%\tikzstyle{relation} = [ellipse, draw, text width=5em, text centered]
+%\tikzstyle{branch} = [rectangle, draw, text width=6em, text centered]
+
 %\usepackage{subfigure}
@@
   \end{lstlisting}
-
+\label{fig:firewallcode}
 \caption{A Stateful Firewall}
@@
 \caption{Signature of the packet portion of the API}
+\end{figure}
+
+  For the stateful firewall included in figure~\ref{fig:firewallcode}, the following decision tree would be generated:
+
+  \begin{figure}
+\label{fig:decisiontree}  
+\caption{Decision Tree for Stateful Firewall}     
 \end{figure}
@@
 
+\subsection*{A.P.I.}
+
+\subsection*{Tracing the Algorithmic Policy}
+From an algorithmic policy, we generate a decision tree  whose leaves are forwarding decision and whose node are either branching according the range of a field value, according to a relation ,or adding and removing components from a relation.
+
+\begin{lstlisting}
+  compile (f: packet -> forwarding_decision) : policy =
+      push most general packet to (empty) stack
+      set the current location to be the root of an empty decision tree
+      while (stack is not empty)
+         pop the top of the stack as pkt
+         run f of pkt resulting in a forwarding decision fd (and a modified packet pkt)
+         record the forwarding decision for the symbolic packet pkt at the current location in the decision tree
+         set the current location to be the root of the decision tree and repeat
+
+\end{lstlisting}
+
+To do so, we repeatedly run the algorithmic policy \lstinline|f| on symbol packets generated to explore every reachable path.
+We begin by calling \lstinline|f| on a symbolic packet covering the full range of value for each of its fields.
+
+When \lstinline|f| calls \lstinline|in_range| to learn if the field \lstinline|fld|'s value is in the range \lstinline|r|, we (1) create an \lstinline|Inrange| branching node in the decision tree (2) refine the current symbolic packet to represent all packets in range \lstinline|r| that could reach this test in the algorithmic policy and return true (3) push on the stack of packets to process a symbolic packets representing all packets in the complement of \lstinline|r| for field \lstinline|fld|.
+
+Similarly, when the algorithmic policy calls \lstinline|in_relation| to learn if fields \lstinline|flds|'s value are in a relation \lstinline|rel|, we (1) create an \lstinline|Inrelation| branching node in the decision tree (2) tag the current packet as being in the relation, return true (3) push on the stack of packets to process the same packet tagged as not being part of relation \lstinline|rel|.
+
+When \lstinline|f| calls \lstinline|add| or \lstinline|remove|, we record in the decision the current symbolic packet, the relation and the fields being added or removed from the relation before continuing with the execution of \lstinline|f|.
+
+When \lstinline|f| returns with a forwarding decision, we record it, paired with the current symbolic packet, as a leaf node at the current position in the decision tree.
+We then pop the next packet from the stack of packets to process and restart tracing the policy from the root of the decision tree, exploring another path of \lstinline|f|.
+
+\subsection*{Simulating the Network Controller}
+
 \subsection*{Optimizations}

### 2015-05-02 draft.tex
..

@@
 f(pkt):
-  if RO.in_range(pkt, src, 0, 10) then
-     if RO.in_range(pkt, dst, 0, 10) then
-        RO.deliver()
+  if Oak.in_range(pkt, src, 0, 10) then
+     if Oak.in_range pkt dst  [(0,10)] then
+        Oak.deliver
      else 
-        RO.add(pkt, (dst, src), trusted);
-        RO.deliver()
-  else if (RO.in_set(pkt, (src,dst), trusted) then
-         RO.deliver()
+        Oak.add pkt (Oak.IpDst, Oak.IpSrc) ``conn'';
+        Oak.deliver
+  else if Oak.in_rel pkt (Oak.IpSrc,Oak.IpDst) ``conn'' then
+         Oak.deliver
        else
@@
    % Why range instead of actual value, pre-generation of rules using symbolic pre-execution
-
-   \begin{figure}
-\begin{lstlisting}
-Module RO
-
-Type pkt (* contains a range for each field *)
-
-Type field (* enum each field *)
-
-function InRange: pkt * field * value * value -> bool
-
-\end{lstlisting}
-
-\caption{Signature of the packet portion of the API}
-\end{figure}
 
@@
 
+\begin{figure}
+  \begin{lstlisting}
+(* a standard packet field *)
+type field = IpSrc | IpDst
+
+(* range is a union of intervals *)
+type range = (int*int) list
+
+(* forwarding decision can deliver, drop, forward, multicast *)
+type forwarding_decision =
+	| Deliver
+	| Drop
+	| ForwardTo of int
+	| Multicast of forwarding_decision * forwarding_decision
+	| Ctrl
+
+(* abstract packet type *)
+type packet
+
+(* minimum, maximum value for a range interval *)
+val min_value: int 
+val max_value: int
+
+(* in_range p f r returns true if the value of field f of packet p is in range r, and false otherwise *)
+val in_range: packet -> field -> range -> bool
+\end{lstlisting}
+
+\caption{Signature of the packet and range-branching portion of the API}
+\end{figure}
+
+\begin{figure}
+  \begin{lstlisting}
+(* name of a relation *)
+type relation = string
+
+(* add packet's fields to a relation *)
+val add: packet -> field list -> relation -> unit
+
+(* remove packet's fields from a relation *)
+val remove: packet -> field list -> relation -> unit
+
+(* test if a packet's fields are in a relation *)
+val in_relation: packet -> field list -> relation -> bool
+\end{lstlisting}
+
+\caption{Signature of the relation portion of the API}
+\end{figure}
+
+\begin{figure}
+  \begin{lstlisting}
+(* compile a user-function into a policy *)
+val compile: (packet -> forwarding_decision) -> policy
+
+(* run the user-defined policy with a set of input packets *)
+val run: policy -> ((field * int) list) list  -> unit
+
+(* Print the policy to stdout *)
+val print_policy: policy -> unit
+  \end{lstlisting}
+  \caption{Signature of the interactive portion of the API}
+\end{figure}
+
 \subsection*{Tracing the Algorithmic Policy}
@@
 
+\begin{figure}
 \begin{lstlisting}
-  compile (f: packet -> forwarding_decision) : policy =
+let compile (f: packet -> forwarding_decision) : policy =
       push most general packet to (empty) stack
@@
 \end{lstlisting}
+\caption{Pseudocode for the compile function}
+\end{figure}
 
@@
 \subsection*{Simulating the Network Controller}
+
+\begin{figure}
+\begin{lstlisting}
+let run (pol: policy) (inputs: ((field*int) list) list) : unit =
+      build the symbolic rules for policy pol
+      generate the concrete rules with all relations empty
+      for each packets in input,
+          find the first of the current rule matching the packet, and record the forwarding decision fd
+          if packet is sent to the controller then
+              evaluate the policy on the packet, potentially adding or removing tuples from relations, record forwarding decision fd'
+              regenerate concrete rules
+              forward the packet according the forwarding decision fd'
+          else
+              forward the packet according to fd 
+\end{lstlisting}
+
+\caption{Pseudocode for the run function}
+  \end{figure}
 

### 2015-05-03 draft.tex
..

@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 \usepackage[latin1]{inputenc}
@@
 \subsection*{A.P.I.}
+In our model, algorithmic policies interact with packets through A.P.I function calls before returning a forwarding decision. We include in Figure~\ref{fig:range_api} the core of our \lstinline|API|.
+
+\lstinline|field| enumerates the different fields of a packet.
+It can be extended to represent more fields of a packet's header.
+Associated with \lstinline|field| are \lstinline|range|s, possible values taken by a certain packet's field, within a given minimum value \lstinline|min_value| and maximum value \lstinline|max_value|.
+\lstinline|range| consists of list of integer pairs, representing a collection of intervals whose union correspond to the range of values of a packet.
+For simplicity, we consider integer fields here, but this could be extended, for example, to IP prefix, boolean flags, etc.
+
+\lstinline|forwarding_decision| enumerates the different forwarding decision which can be returned by an algorithmic policy.
+\lstinline|Deliver| forwards the packet to its destination, \lstinline|Drop| drops the packet, \lstinline|ForwardTo i| forwards the packet to a constant location \lstinline|i| and \lstinline|Multicast fd1 fd2| duplicates the packet and handles each according to \lstinline|fd1| and \lstinline|fd2| respectively.
 
@@
   \begin{lstlisting}
+(* abstract packet type *)
+type packet
+
 (* a standard packet field *)
@@
 type range = (int*int) list
+
+(* minimum, maximum value for a range interval *)
+val min_value: int 
+val max_value: int
 
@@
 	| Multicast of forwarding_decision * forwarding_decision
-	| Ctrl
-
-(* abstract packet type *)
-type packet
-
-(* minimum, maximum value for a range interval *)
-val min_value: int 
-val max_value: int
+	| Ctrl (* TODO: should not expose this *)
 
@@
 \end{lstlisting}
-
+\label{fig:range_api}
 \caption{Signature of the packet and range-branching portion of the API}
@@
 \end{lstlisting}
-
+\label{fig:rel_api}
 \caption{Signature of the relation portion of the API}
@@
   \end{lstlisting}
+  \label{fig:build_api}
   \caption{Signature of the interactive portion of the API}

### 2015-05-03 draft.tex
..

@@
 
-% Including an example policy, todo change 0 to 10 to actual IP
 Figure~\ref{fig:firewall-eg} displays a policy for a stateful firewall using our API.
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \begin{lstlisting}
@@
   \end{lstlisting}
+
+\caption{A Stateful Firewall}
 \label{fig:firewallcode}
-\caption{A Stateful Firewall}
 \end{figure}
@@
   \begin{figure}
+
+\caption{Decision Tree for Stateful Firewall}     
 \label{fig:decisiontree}  
-\caption{Decision Tree for Stateful Firewall}     
   \end{figure}
@@
 
-   \begin{figure}
-     \begin{lstlisting}
-
-function InRange: pkt * field * value * value -> bool
-
-\end{lstlisting}
-
-\caption{Signature of the relation portion of the API}
-\end{figure}
-
    % Computing which packets need to be sent to the controller (i.e. only the ones reaching an ``add'' or a ``remove'' node.
@@
 \subsection*{A.P.I.}
-In our model, algorithmic policies interact with packets through A.P.I function calls before returning a forwarding decision. We include in Figure~\ref{fig:range_api} the core of our \lstinline|API|.
-
+In our model, algorithmic policies interact with packets through A.P.I function calls before returning a forwarding decision.
+We include in Figure~\ref{fig:range_api} the core of our \lstinline|A.P.I.|.
 \lstinline|field| enumerates the different fields of a packet.
@@
 
-\begin{figure}
+Algorithmic policy may base their forwarding decision on the result of calling the \lstinline|in_range pkt fld r|, which returns whether value of packet \lstinline|pkt| at field \lstinline|fld| is in the range \lstinline|r|.
+
+\begin{figure}[ht]
   \begin{lstlisting}
@@
 \end{lstlisting}
+
+\caption{Signature of the packet and range-branching portion of the API}
 \label{fig:range_api}
-\caption{Signature of the packet and range-branching portion of the API}
 \end{figure}
 
-\begin{figure}
+In Figure~\ref{fig:rel_api}, we include the relation portion of the \lstinline|A.P.I.|, used to model stateful controller programs.
+A \lstinline|relation| represents a table in which each row contains a series of field values.
+It is referred to by a string.
+Policy writer may add a row to a relation \lstinline|rel| by calling the function \lstinline|add pkt flds rel|, which will populate the new row with the fields \lstinline|flds|' values of the current packet \lstinline|pkt|.
+Similarly, they may remove a row containing exactly the values of fields \lstinline|flds|' values of the current packet \lstinline|pkt| by calling \lstinline|remove pkt flds rel|.
+Finally, one may test if the values of fields \lstinline|flds| of current packet \lstinline|pkt| are in relation \lstinline|rel| by calling \lstinline|in_relation pkt flds rel|.
+
+\begin{figure}[ht]
   \begin{lstlisting}
@@
 \end{lstlisting}
+
+\caption{Signature of the relation portion of the API}
 \label{fig:rel_api}
-\caption{Signature of the relation portion of the API}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[ht]
   \begin{lstlisting}
@@
   \end{lstlisting}
+
+  \caption{Signature of the interactive portion of the API}
   \label{fig:build_api}
-  \caption{Signature of the interactive portion of the API}
 \end{figure}
@@
 
-When \lstinline|f| calls \lstinline|in_range| to learn if the field \lstinline|fld|'s value is in the range \lstinline|r|, we (1) create an \lstinline|Inrange| branching node in the decision tree (2) refine the current symbolic packet to represent all packets in range \lstinline|r| that could reach this test in the algorithmic policy and return true (3) push on the stack of packets to process a symbolic packets representing all packets in the complement of \lstinline|r| for field \lstinline|fld|.
+When \lstinline|f| calls \lstinline|in_range| to learn if the field \lstinline|fld|'s value is in the range \lstinline|r|, we
+\begin{enumerate}
+\item  create an \lstinline|Inrange| branching node in the decision tree
+\item refine the current symbolic packet to represent all packets in range \lstinline|r| that could reach this test in the algorithmic policy and return true
+\item push on the stack of packets to process a symbolic packets representing all packets in the complement of \lstinline|r| for field \lstinline|fld|.
+\end{enumerate}
 
-Similarly, when the algorithmic policy calls \lstinline|in_relation| to learn if fields \lstinline|flds|'s value are in a relation \lstinline|rel|, we (1) create an \lstinline|Inrelation| branching node in the decision tree (2) tag the current packet as being in the relation, return true (3) push on the stack of packets to process the same packet tagged as not being part of relation \lstinline|rel|.
-
+Similarly, when the algorithmic policy calls \lstinline|in_relation| to learn if fields \lstinline|flds|'s value are in a relation \lstinline|rel|, we
+\begin{enumerate}
+  \item create an \lstinline|Inrelation| branching node in the decision tree
+  \item tag the current packet as being in the relation, return true
+  \item push on the stack of packets to process the same packet tagged as not being part of relation \lstinline|rel|.
+\end{enumerate}
 When \lstinline|f| calls \lstinline|add| or \lstinline|remove|, we record in the decision the current symbolic packet, the relation and the fields being added or removed from the relation before continuing with the execution of \lstinline|f|.

### 2015-05-05 draft.tex
..
two column

@@
-\documentclass[11pt]{article}
+\documentclass[preprint]{sigplanconf}
 \usepackage[latin1]{inputenc}
-
 \usepackage{times}
 \usepackage{latexsym}
-\usepackage[bottom=20mm, top=20mm, left=20mm, right=20mm]{geometry}
 \usepackage{graphicx}
@@
 
-\section*{API and Decision Trees}
+\section*{Decision Trees and Algorithmic Policies}
    \subsection*{Range-based Branching}
@@
    % Computing which packets need to be sent to the controller (i.e. only the ones reaching an ``add'' or a ``remove'' node.
+
+   \subsection*{Combining Algorithmic Policies}
+   % cross, ``holes''
 
@@
 \caption{Pseudocode for the compile function}
+\label{fig:compile-pseudo}
 \end{figure}
@@
 
+We provide a function \lstinline|run| to simulate the controller policy on a sequence of packets.
+\lstinline|policy| is an abstract datatype solely returned by \lstinline|compile| (see Fig.~\ref{fig:compile-pseudo}).
+In addition to the policy, \lstinline|run| takes in a list of packets represented as list of pairs of \lstinline|field| and \lstinline|int|, representing the values of the different fields of the packet.
+We expect all fields to be set exactly once for each packet in the list.
+
+\lstinline|run| starts by building the symbolic rules for the given policy, and generating the concrete rules assuming all relations to be empty.
+In sequence, packet are matched with the current concrete rules, recording the forwarding decision.
+If a packet is sent to the controller, we evaluate the policy on the packet, potentially adding or removing tuples from relations.
+We then re-generate the concrete rules with the updated relation, before recording the forwarding decision.
+
 \begin{figure}[ht]
@@
 \caption{Pseudocode for the run function}
+\label{fig:run-pseudo}
   \end{figure}
@@
 
+% probably no evaluation :(
 \section*{Evaluation}

### 2015-05-05 draft.tex
..

@@
 
+\title{Red Oak: Something Something Algorithmic Policy}
+ \authorinfo{Ryan Beckett}{Princeton University}{rbeckett@princeton.edu}
+ \authorinfo{Olivier Savary B\'{e}langer}{Princeton University}{olivierb@princeton.edu}
+
 %-----------------------------------------------------------
@@
 \begin{document}
-\quad \\
-\quad \\
-\quad \\
-\centerline{ \large{Red Oak: Something Something Algorithmic Policy}}
-\vspace{1pt}\\
-\centerline{Ryan Beckett and Olivier Savary B\'{e}langer} \\
-\\
-\centerline{April 15, 2015} \\
+\maketitle
 

### 2015-05-05 draft.tex
..

@@
 
+\begin{abstract} ...
+\end{abstract}
+
 \section*{Introduction}
@@
 f(pkt):
-  if Oak.in_range(pkt, src, 0, 10) then
-     if Oak.in_range pkt dst  [(0,10)] then
-        Oak.deliver
+  if in_range(pkt, src, 0, 10) then
+     if in_range pkt dst  [(0,10)] then
+        deliver
      else 
-        Oak.add pkt (Oak.IpDst, Oak.IpSrc) ``conn'';
-        Oak.deliver
-  else if Oak.in_rel pkt (Oak.IpSrc,Oak.IpDst) ``conn'' then
-         Oak.deliver
+        add pkt (IpDst, IpSrc) ``conn'';
+        deliver
+  else if in_rel pkt (IpSrc,IpDst) ``conn'' then
+         deliver
        else

### 2015-05-06 draft.tex
..

@@
 \section*{Decision Trees and Algorithmic Policies}
+
    \subsection*{Range-based Branching}
@@
 
-   \subsection*{Forwarding Decision}
-   % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to dst)
-
      \subsection*{Built-in Relation with Automatic Invalidation}
@@
 
+   \subsection*{Forwarding Decision}
+   % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to dst)
+   Under the big switch model, the forwarding decisions returned by algorithmic policies are destination-based rather than path-based.
+For example, the algorithmic policy may decide to deliver a packet, at which point this interpretation will be interpreted as forwarding a packet along a path terminating at the packet's destination.
+Packets may also be dropped, or forwarding to a static location.
+Our model also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller.
+
    \subsection*{Combining Algorithmic Policies}
    % cross, ``holes''
+   As algorithmic policies under our model are functions of the host language, we inherit the latter's function space ability, allowing, amongst other things, for compositon of algorithmic policies and higher-order algorithmic policies.
+
+\begin{figure}[ht]
+\begin{lstlisting}
+
+let filter_src_range  (f: packet -> forwarding_decision) (r:range) (pkt: packet) : forwarding_decision =
+ if in_range pkt IpSrc r then
+    f pkt
+ else
+   Drop   
+\end{lstlisting}
+
+\caption{Example of a Partial Algorithmic Policy}
+\label{fig:ex-hole}
+\end{figure}
+
+   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_src_range| a partial algorithmic policy which, given an algorithmic policy \lstinline|f| and a range \lstinline|r|, will forwards all packets whose source IP is within range \lstinline|r| and drop the packet otherwise.
+
+   \begin{figure}[ht]
+\begin{lstlisting}
+let f_cross (f_1) (f_2) pkt =
+  let d_1 = f_1 pkt in
+  let d_2 = f_2 pkt in
+  Multicast (d_1,d_2)
+\end{lstlisting}
+
+\caption{Example of a Composition Operator}
+\label{fig:ex-cross}
+\end{figure}
+
+Various composition operator may be built as host language function to combine multiple algorithmic policies into a single one.
+For example, \lstinline|f_cross|, included in Figure~\ref{fig:ex-cross}, runs algorithmic policy \lstinline|f_1| and \lstinline|f_2| on input packets and multicast them according to both forwarding decisions.
 

### 2015-05-06 draft.tex
..

@@
 
-let filter_src_range  (f: packet -> forwarding_decision) (r:range) (pkt: packet) : forwarding_decision =
+  let filter_src_range
+  (f: packet -> forwarding_decision)
+  (r:range)
+  (pkt: packet) : forwarding_decision =
  if in_range pkt IpSrc r then
@@
   \begin{lstlisting}
-(* abstract packet type *)
 type packet
 
-(* a standard packet field *)
 type field = IpSrc | IpDst
 
-(* range is a union of intervals *)
 type range = (int*int) list
 
-(* minimum, maximum value for a range interval *)
 val min_value: int 
@@
 
-(* forwarding decision can deliver, drop, forward, multicast *)
 type forwarding_decision =
@@
 	| Multicast of forwarding_decision * forwarding_decision
-	| Ctrl (* TODO: should not expose this *)
+	| Ctrl
 
-(* in_range p f r returns true if the value of field f of packet p is in range r, and false otherwise *)
 val in_range: packet -> field -> range -> bool
@@
   \begin{lstlisting}
-(* name of a relation *)
 type relation = string
 
-(* add packet's fields to a relation *)
 val add: packet -> field list -> relation -> unit
 
-(* remove packet's fields from a relation *)
 val remove: packet -> field list -> relation -> unit
 
-(* test if a packet's fields are in a relation *)
 val in_relation: packet -> field list -> relation -> bool
@@
 
-\begin{figure}[ht]
+\begin{figure*}[ht]
 \begin{lstlisting}
@@
 \label{fig:compile-pseudo}
-\end{figure}
+\end{figure*}
 
@@
 \label{fig:run-pseudo}
-  \end{figure}
+  \end{figure*}
 

### 2015-05-06 draft.tex
removed Ctrl from forwarding_policy, instead representing rules as forwarding_decision option with None interpreted as Ctrl
..

@@
 Packets may also be dropped, or forwarding to a static location.
-Our model also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller.
+Our model also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller (not shown in Fig~\ref{fig:range_api}).
 
@@
 
-  let filter_src_range
-  (f: packet -> forwarding_decision)
-  (r:range)
-  (pkt: packet) : forwarding_decision =
+let filter_src_range f r pkt =
  if in_range pkt IpSrc r then
@@
 	| Multicast of forwarding_decision * forwarding_decision
-	| Ctrl
 

### 2015-05-08 draft.tex
..

@@
        else
-         RO.drop
+         drop
        end
@@
 
-\section*{Decision Trees and Algorithmic Policies}
-
-   \subsection*{Range-based Branching}
-   % Why range instead of actual value, pre-generation of rules using symbolic pre-execution
-
+From an algorithmic policy such as the one included in Fig~\ref{fig:firewall-eg}, we generate a decision tree summarizing the function's interaction with the packet and the relevant decision.
+The decision tree is then used by our runtime system to install switch rules and update them when necessary.
 For the stateful firewall included in figure~\ref{fig:firewallcode}, the following decision tree would be generated:
 
-  \begin{figure}
+  \begin{figure}[ht]
 
@@
   \end{figure}
+
+  This decision tree would in turn lead to the following symbolic rules:
+  \begin{lstlisting}
+ipsrc in [0, 100] --> Sent to Controller
+ipsrc in [101, 1000], (ipsrc,ipdst) in conn
+                  --> Deliver
+ipsrc in [101, 1000], (ipsrc,ipdst) not in conn
+                  --> Drop
+  \end{lstlisting}
+
+  At initialization, these concrete rules would be generated from the symbolic rules by having all relations be empty:
+\begin{lstlisting}
+ipsrc in [0, 100] ---> Sent to Controller  
+ipsrc in [101, 1000] ---> Drop  
+\end{lstlisting}
+
+  When a packet is received, it is forwarded according to the first matching concrete rule installed on the switch.
+If the packet is sent to the controller, we evaluate the decision tree for that packet, potentially adding and removing tuples from relations, before forwarding the packet according to forwarding decision held in the relevant leaf of the decision tree.
+We then regenerate the concrete rules using the updated relation.
+For example, after having seen packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
+
+\begin{lstlisting}
+ipsrc in [0, 100] ---> Sent to Controller  
+ipdst = 10, ipsrc = 150 ---> Deliver  
+ipsrc in [101, 1000] ---> Drop  
+\end{lstlisting}
+
+\section*{Algorithmic Policies and Decision Trees}
+In this section, we describe how an algorithmic policy may interact with a packet and with relations to generate a forwarding decision, and how we build a decision tree out of an algorithmic policy.
+
+   \subsection*{Range-based Branching}
+
+   % Why range instead of actual value, pre-generation of rules using symbolic pre-execution
 

### 2015-05-08 draft.tex
lstlisting keyword for api

@@
 %\subsectionfont{\fontsize{11}{14}\selectfont}
+
+\lstset{morekeywords={in_range, in_rel, add, remove, Drop, Deliver, ForwardTo, Multicast, IpSrc, IpDst}}
 
@@
      if in_range pkt dst  [(0,10)] then
-        deliver
+        Deliver
      else 
         add pkt (IpDst, IpSrc) ``conn'';
-        deliver
+        Deliver
   else if in_rel pkt (IpSrc,IpDst) ``conn'' then
-         deliver
+         Deliver
        else
-         drop
+         Drop
        end
@@
 Similarly, they may remove a row containing exactly the values of fields \lstinline|flds|' values of the current packet \lstinline|pkt| by calling \lstinline|remove pkt flds rel|.
-Finally, one may test if the values of fields \lstinline|flds| of current packet \lstinline|pkt| are in relation \lstinline|rel| by calling \lstinline|in_relation pkt flds rel|.
+Finally, one may test if the values of fields \lstinline|flds| of current packet \lstinline|pkt| are in relation \lstinline|rel| by calling \lstinline|in_rel pkt flds rel|.
 
@@
 
-val add: packet -> field list -> relation -> unit
+val add: packet -> field list -> relation
+    -> unit
 
-val remove: packet -> field list -> relation -> unit
+val remove: packet -> field list -> relation
+    -> unit
 
-val in_relation: packet -> field list -> relation -> bool
+val in_rel: packet -> field list -> relation
+    -> bool
 \end{lstlisting}
@@
   \begin{lstlisting}
-(* compile a user-function into a policy *)
-val compile: (packet -> forwarding_decision) -> policy
+val compile: (packet -> forwarding_decision)
+    -> policy
 
-(* run the user-defined policy with a set of input packets *)
-val run: policy -> ((field * int) list) list  -> unit
+val run: policy -> ((field * int) list) list
+    -> unit
 
-(* Print the policy to stdout *)
 val print_policy: policy -> unit

### 2015-05-09 draft.tex
intro

@@
 
-\title{Red Oak: Something Something Algorithmic Policy}
+\title{Red Oak: Proactive Policy Generation}
  \authorinfo{Ryan Beckett}{Princeton University}{rbeckett@princeton.edu}
@@
 
-Maple \cite{Maple} proposes a SDN controller model where the behavior of the controller is defined by a function from packets to forwarding path, the algorithmic policy.
-The authors then show how to optimize the controller so that packets are handled without having to run the function on them.
+One of the major outcomes of the Software Defined Networking (SDN) movement, has been that of a unified, open interface to the switch hardware.
+This has, in turn, enabled new opportunities for network programmability.
+The typical model for SDN is that of a centralized controller, or distributed set of controllers, programming the switches dynamically to implement the desired control functionality.
+However the questions remains as to how best to write programs for SDNs.
+
+Recent collaborations between researches in networking and programming languages have resulting in a number of new, network-specific programming languages.
+One example is the Frenetic family of languages \cite{Frenetic}, which allow programmers to write simple packet processing functions in a domain-specific language and combine them together using policy combinators.
+Another approach given in Maple \cite{Maple} is to allow programmers to write policies in an arbitrary, high-level programming language as a function that takes a packet and produces a forwarding path, the algorithmic policy.
+Conceptually, the function in run anew on each packet that enters the network.
+
+Ultimately, the performance requirements of networks dictate that the centralized controller(s) can not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency.
+For this reason, compiler technologies for network programming languages must be able to push much of the control plane functionality into the data plane switches themselves.
+
+For Maple, the authors show how to optimize the controller so that packets are handled without having to run the function on them.
 Maple does so by providing a packet API capturing each access to packet fields, and memoizing the results of the function for subsequent packets sharing the accessed fields' values.
-We identify a number of shortcoming to this model:
+Unfortunately, there are a number of limitations to this approach:
+
 \begin{enumerate}
-\item When a field value is observe, the forwarding decision only works on packets with this exact value, even if the forwarding decision is made on a much coarser range of value.Similarly, after observing some value in the environment, the forwarding decision is specialize for that particular value
-
-\item It no built-in environment datastructure with automatic invalidation of decision trace when the environment is modified, which is needed to implement dynamic controllers.
-
-  \item Populating the decision tree requires seeing a high number of packets, potentially slowing down the network at start-up. 
+\item Since controller decisions are memoized, all initial packets must first go to the controller before it can learn how to forward similar packets directly in the data plane.
+This leads to a cold-startup phase with high packet latency.
+\item The controller only observes concrete packets.
+This can lead to installing overly specialized forwarding rules on the switches that miss some of the high-level intent of the programmer (e.g., matching all packets with a particular IP prefix).
+\item Most interesting network functions keep some global state (e.g., MAC learning table, ACLs).
+To ensure memoized decisions do not become invalid when the controller's environment (global state) changes, programmers must manually decide which previous decisions to invalidate.
 \end{enumerate}
 
-To address these, we modify the model in the following way:
+It no built-in environment datastructure with automatic invalidation of decision trace when the environment is modified, which is needed to implement dynamic controllers.
+
+In this work, we language-independent API and compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
+However, we address the shortcomings in Maple in the following way:
+
 \begin{enumerate}
-  \item calls to our API does not expose the values of a packet's field, answering instead boolean queries about the values being in a certain range
-  \item dynamic controllers are supported through a relation data-structure which is updated through an extended API
-  \item the trace tree can be generated from executing the policy on a single (potentially symbolic) packet.
+\item We describe a static analysis on the user-defined forwarding policy to proactively generate forwarding rules to install in the network.
+\item We provide a packet API that lets programmers ask if a field is in a particular range (e.g., IP prefix) rather than if the the field equals a particular value.
+This more expressive test works together with our analysis to better capture programmer intent when setting up forwarding decision in the data plane.
+\item We provide a relation/table data structure in our API to maintain global state.
+The data structure works together with our analysis to provide an automatic invalidation scheme.
+The analysis determines which future packets might affect global state and ensures that these relevant packets are actually seen by the controller.
 \end{enumerate}

### 2015-05-09 draft.tex
added to example

@@
 
-It no built-in environment datastructure with automatic invalidation of decision trace when the environment is modified, which is needed to implement dynamic controllers.
-
 In this work, we language-independent API and compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
-However, we address the shortcomings in Maple in the following way:
+However, we address each of the shortcomings in Maple listed above. Our contributions are the following:
 
 \begin{enumerate}
-\item We describe a static analysis on the user-defined forwarding policy to proactively generate forwarding rules to install in the network.
+\item We describe a static analysis on the user-defined forwarding policy that pushes symbolic packets through the function to explore all possible paths of execution.
+This analysis is then used to proactively generate forwarding rules to install in the network.
 \item We provide a packet API that lets programmers ask if a field is in a particular range (e.g., IP prefix) rather than if the the field equals a particular value.
@@
 Algorithmic policies were introduced by Maple as a model for SDN controllers.
-An algorithmic policy consist of a function from packet to forwarding decision.
-Under this model, the network forwards packets as if this function was executed on each packet when entering the network.
+An algorithmic policy is a programmer-defined function from a packet to a forwarding decision (i.e., where the packet(s) should end up).
+Conceptually this model offers the abstraction that the network forwards packets as if this function were executed on each packet when entering the network.
+To handle future traffic directly in the data plane, Maple keeps a tree data structure that branches on decisions observed at run time (e.g., the packet had source IP 10.0.0.1)
 
-Figure~\ref{fig:firewall-eg} displays a policy for a stateful firewall using our API.
-Sources from 0 to 10 are taken to be internal to the network.
+Figure~\ref{fig:firewall-eg} displays a policy for a simple stateful firewall using our API.
+For simplicity, we assume that IP ranges are between 0 and 1000, however it is easy to generalize this to actual IP prefixes.
+In the example, source IPs from 0 to 10 are considered to be internal to the network and everything else, external.
+
 If a packet is sent from an internal host to an external host, the pair of internal and external host is added to a list of trusted connection.
-If a packet is received from an external host, we lookup the table to see if the connection is trusted, in which case we forward the packet to its destination, otherwise we drop it.
+If a packet is received from an external host, we lookup in the table to see if the connection is trusted, in which case we forward the packet to its destination, otherwise we drop it.
+The forwarding decision "Deliver" is just shorthand for "forward the packet according to its intended destination".
+An interesting point to be made about this function is that we are actively manipulating global policy state when we add entries to the stateful firewall's trusted connection relation.
 
@@
 
-From an algorithmic policy such as the one included in Fig~\ref{fig:firewall-eg}, we generate a decision tree summarizing the function's interaction with the packet and the relevant decision.
-The decision tree is then used by our runtime system to install switch rules and update them when necessary.
+From an algorithmic policy such as the one included in Fig~\ref{fig:firewall-eg}, we generate a decision tree summarizing the function's interaction with the packet and the relevant decisions that will be made for particular classes of packets.
+The decision tree is created by analyzing the programmers function statically and then used by our runtime system to install switch rules and update them only when necessary (i.e., in response to changes in global state).
 For the stateful firewall included in figure~\ref{fig:firewallcode}, the following decision tree would be generated:
@@
 
-  This decision tree would in turn lead to the following symbolic rules:
+This decision tree would, in turn, lead to the following set of prioritized symbolic rules.
+We call these symbolic rules because they represent constraints on the actual concrete rules.
+The concrete rules our system will install in the network depend on the current contents of the trusted connection relation.
+In particular, for each IP source that is trusted to communicate with an internal destination in the "conn" relation, we will get a concrete rule permitting this communication.
   \begin{lstlisting}
@@
 
-  At initialization, these concrete rules would be generated from the symbolic rules by having all relations be empty:
+After initialization (starting with no trusted connections), the concrete rules would be generated from the symbolic rules by having all relations be empty:
 \begin{lstlisting}
@@
 We then regenerate the concrete rules using the updated relation.
+
  For example, after having seen packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
@@
 ipsrc in [0, 100] ---> Sent to Controller  
-ipdst = 10, ipsrc = 150 ---> Deliver  
+ipsrc = 150, ipdst = 10 ---> Deliver  
 ipsrc in [101, 1000] ---> Drop  

### 2015-05-09 draft.tex
decision tree image placeholder
small pass on the intro

@@
 
-One of the major outcomes of the Software Defined Networking (SDN) movement, has been that of a unified, open interface to the switch hardware.
+One of the major outcomes of the Software Defined Networking (SDN) movement has been a unified, open interface to the switch hardware.
 This has, in turn, enabled new opportunities for network programmability.
-The typical model for SDN is that of a centralized controller, or distributed set of controllers, programming the switches dynamically to implement the desired control functionality.
-However the questions remains as to how best to write programs for SDNs.
+The typical model for SDN involves a centralized controller or distributed set of controllers programming switches dynamically to implement the desired control functionality.
+However, the questions remains as to how best to write programs for SDNs.
 
-Recent collaborations between researches in networking and programming languages have resulting in a number of new, network-specific programming languages.
+Recent collaborations between researchers in networking and programming languages have resulting in a number of new network-specific programming languages.
 One example is the Frenetic family of languages \cite{Frenetic}, which allow programmers to write simple packet processing functions in a domain-specific language and combine them together using policy combinators.
-Another approach given in Maple \cite{Maple} is to allow programmers to write policies in an arbitrary, high-level programming language as a function that takes a packet and produces a forwarding path, the algorithmic policy.
-Conceptually, the function in run anew on each packet that enters the network.
+Another approach, described in Maple \cite{Maple}, is to allow programmers to write policies in an arbitrary, high-level programming language as a function that takes a packet and produces a forwarding path, the algorithmic policy.
+Conceptually, the function is run anew on each packet that enters the network.
 
-Ultimately, the performance requirements of networks dictate that the centralized controller(s) can not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency.
+Ultimately, the performance requirements of networks dictate that the centralized controller(s) can not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency compared to those being handled by switches.
 For this reason, compiler technologies for network programming languages must be able to push much of the control plane functionality into the data plane switches themselves.
@@
 This more expressive test works together with our analysis to better capture programmer intent when setting up forwarding decision in the data plane.
-\item We provide a relation/table data structure in our API to maintain global state.
+\item We provide a mutable relation data structure in our API to maintain global state.
 The data structure works together with our analysis to provide an automatic invalidation scheme.
@@
 \begin{figure}[ht]
-
+\includegraphics[scale=.5]{img/dtree.png}
 \caption{Decision Tree for Stateful Firewall}     
@@
    % Can represent constraints on the ranges admissible given a certain topology as a series of ``true'' branches at the root of the tree.
+im
 

### 2015-05-09 draft.tex
..

@@
 
-Figure~\ref{fig:firewall-eg} displays a policy for a simple stateful firewall using our API.
+A policy for a simple stateful firewall using our API is included in Figure~\ref{fig:firewallcode}.
 For simplicity, we assume that IP ranges are between 0 and 1000, however it is easy to generalize this to actual IP prefixes.
@@
 
- For example, after having seen packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
+ For example, after having seen a packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
 
@@
 	| ForwardTo of int
-	| Multicast of forwarding_decision * forwarding_decision
+	| Multicast of forwarding_decision
+                     l* forwarding_decision
 
@@
    % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
+The rules generated by the presented technique ensures that the only packets being forwarded to the controller are those reaching an add or a remove node and thus potentially modify the state of the controller.
+However, these packets do not always modify the state of the controller.
+For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|ipsrc| of 10 and \lstinline|ipdst| of 150, any subsequent packet with the same \lstinline|ipsrc| and \lstinline|ipdst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent.
+By using both local and global transformations on the decision tree, we can get closer to an optimal amount of packet forwarded to the controller, where only packets modifying relations are seen by the controller.
+
+We implemented in Red Oak, under \lstinline|o1|, a local transformation adding an \lstinline|in_rel| node before \lstinline|add| or \lstinline|remove| nodes.
+In the case of \lstinline|add|, the \lstinline|add| node is omited in the true branch of \lstinline|in_rel|, such that packets whose fields are already in the relation will not be seen by the controller due to this \lstinline|add| node.
+Similarly, we omit the \lstinline|remove| node in the false branch of \lstinline|in_rel|.
+In the firewall example (see Fig.~\ref{fig:firewallcode}), this would result the following decision tree:
+
+\begin{figure}[ht]
+  TODO
+\caption{Optimized Decision Tree for Stateful Firewall}     
+\label{fig:decisiontreeopt}  
+  \end{figure}
+
+From which the following symbolic rule would be generate:
+\begin{lstlisting}
+ipsrc in [0, 100], (ipdst, ipsrc) in conn
+                  --> Deliver
+ipsrc in [0, 100] --> Sent to Controller
+ipsrc in [101, 1000], (ipsrc,ipdst) in conn
+                  --> Deliver
+ipsrc in [101, 1000], (ipsrc,ipdst) not in conn
+                  --> Drop
+\end{lstlisting}
+
+With empty relation, the concrete rules would be the same as with the unoptimized decision tree.
+However, after seeing, for example, packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the rules would be
+\begin{lstlisting}
+ipsrc = 10, ipdst = 150 ---> Deliver  
+ipsrc in [0, 100] ---> Sent to Controller  
+ipsrc = 150, ipdst = 10 ---> Deliver  
+ipsrc in [101, 1000] ---> Drop  
+\end{lstlisting}
+
    % 2) Prune dead subtree on empty ranges and add/remove which do not affect any inhabitant-based branching.

### 2015-05-09 draft.tex
opt

@@
 
+% Prune dead subtree on empty ranges and add/remove which do not affect any inhabitant-based branching.
+
 When \lstinline|f| returns with a forwarding decision, we record it, paired with the current symbolic packet, as a leaf node at the current position in the decision tree.
@@
 
-   % 2) Prune dead subtree on empty ranges and add/remove which do not affect any inhabitant-based branching.
+Other optimizations may further reduce the number of packet sent to the controller.
+For example, we could remove \lstinline|add| node which are followed by \lstinline|remove| node for the same relation and tuple on all path without the change being observed by an \lstinline|in_rel| node.
+Similarly, we could detect and warn the user if an added tuple may never be observed by an \lstinline|in_rel| node, for example by switching \lstinline|ipsrc| and \lstinline|ipdst| in the add tuple of the stateful firewall example (see Figure~\ref{fig:firewallcode}).
+We leave the implementation of these optimizations as future work.
 

### 2015-05-09 draft.tex
range based

@@
    % Why range instead of actual value, pre-generation of rules using symbolic pre-execution
+   Our API exports a single function for inspecting the contents of a packet's field, which is a test on the range of the packet's field.
+The ability to test if a value falls in a particular range is useful for capturing programmer intent with respect to things like IP prefix matching.
+Other tests on packet fields can be encoded in terms of ranges.
+For example, to test whether a field equals a particular value, one can simply test if it falls in the range with that value as both the lower and upper bound.
+Likewise, the programmer can test if a field is greater than or less than a particular value by testing if the field is in the range from the value to the maximum or minimum possible value for that field respectively.
+
+Conveniently, ranges on packet fields also give us a way to describe symbolic packets - packets that represent a set of concrete packets.
+Later we will see how our static analysis will proactively generate rules to install on the switch hardware by propagating symbolic packet ranges through the control structure of the program.
 
    % Can represent constraints on the ranges admissible given a certain topology as a series of ``true'' branches at the root of the tree.
-im
 

### 2015-05-09 draft.tex
built-in relations

@@
    % Relation with add and remove tuple of fields from a relation (empty at the begining), Inhabitant-based Branching 
+	Many network applications of interest need to keep around some global state to keep track of things that have happened in the network.
+A stateful firewall, like the one we introduced earlier, needs to maintain an ACL of external hosts that are trusted to communicate with internal hosts.
+This list can change over time as packets traverse the network.
+Likewise, a MAC learning application would need to remember where to send packets destined for particular Ethernet destinations.
+However, reasoning about which packets the controller needs to see, and how to set up forwarding rules in the data plane to ensure that these packets are actually seen by the controller is challenging for programmers.
+
+	We introduce a simple table (relation) data structure that handles updating mutable, global state in the controller automatically.
+Intuitively, each time a table can be updated, the controller must witness the update.
+Our range-based analysis determines all possible packets that could reach code resulting in an update to the table and instruments the data plane with forwarding rules to send relevant traffic to the controller.
 

### 2015-05-09 draft.tex
forwarding decisions

@@
    % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to dst)
-   Under the big switch model, the forwarding decisions returned by algorithmic policies are destination-based rather than path-based.
-For example, the algorithmic policy may decide to deliver a packet, at which point this interpretation will be interpreted as forwarding a packet along a path terminating at the packet's destination.
-Packets may also be dropped, or forwarding to a static location.
+	The programming model for Oak is that each packet entering the network is mapped to a forwarding decision that describes where the packet should end up leaving the network.
+For simplicity, we assume that we are targeting a one big switch implementation of the network.
+The one big switch model has been heavily optimized \cite{Obs}, and we view the problem of determining paths through the network as orthogonal to our work here.
+
+   We provide several types of forwarding decisions to the programmer.
+The programmer can deliver a packet, according to the packet's desination, drop a packet, and forwarding a packet to a static location.
 Our model also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller (not shown in Fig~\ref{fig:range_api}).
+Finally, the programmer can combine several forwarding decisions into one by using a mulitcast forwarding decision.
 

### 2015-05-09 draft.tex
combinators

@@
 
-   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_src_range| a partial algorithmic policy which, given an algorithmic policy \lstinline|f| and a range \lstinline|r|, will forwards all packets whose source IP is within range \lstinline|r| and drop the packet otherwise.
+   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_src_range| a partial policy, which, given a policy \lstinline|f| and a range \lstinline|r|, will forwards all packets whose source IP is within range \lstinline|r| and drop the packet otherwise.
+The ability to compose two or more policies in sequence like this arises many times in practice.
+In our earlier firewall example, instead of delivering packets that were not dropped, we could instead have deferred to another function to decide how to handle trusted traffic.
 
@@
 Various composition operator may be built as host language function to combine multiple algorithmic policies into a single one.
-For example, \lstinline|f_cross|, included in Figure~\ref{fig:ex-cross}, runs algorithmic policy \lstinline|f_1| and \lstinline|f_2| on input packets and multicast them according to both forwarding decisions.
+For example, \lstinline|f_cross|, included in Figure~\ref{fig:ex-cross}, runs algorithmic policy \lstinline|f_1| and \lstinline|f_2| on input packets and multicasts them according to both forwarding decisions.
+This corresponds to the parallel composition operator for the Frenetic language \cite{Frenetic}, however here the composition takes place in a fully general purpose language.
 

### 2015-05-09 draft.tex
minor changes

@@
 	| Multicast of forwarding_decision
-                     l* forwarding_decision
+                      * forwarding_decision
 
@@
 
-val add: packet -> field list -> relation
-    -> unit
+val add: packet -> field list -> relation -> unit
 
-val remove: packet -> field list -> relation
-    -> unit
+val remove: packet -> field list -> relation -> unit
 
-val in_rel: packet -> field list -> relation
-    -> bool
+val in_rel: packet -> field list -> relation-> bool
 \end{lstlisting}

### 2015-05-10 draft.tex
..

@@
 
-\title{Red Oak: Proactive Policy Generation}
+\title{\textit{Red Oak}: Proactive Policy Generation}
  \authorinfo{Ryan Beckett}{Princeton University}{rbeckett@princeton.edu}
@@
 
-val add: packet -> field list -> relation -> unit
+val add: packet -> field list -> relation
+    -> unit
 
-val remove: packet -> field list -> relation -> unit
+val remove: packet -> field list -> relation
+    -> unit
 
-val in_rel: packet -> field list -> relation-> bool
+val in_rel: packet -> field list -> relation
+    -> bool
 \end{lstlisting}
@@
 
-We implemented in Red Oak, under \lstinline|o1|, a local transformation adding an \lstinline|in_rel| node before \lstinline|add| or \lstinline|remove| nodes.
+We implemented in \textit{Red Oak} a local transformation adding an \lstinline|in_rel| node before \lstinline|add| or \lstinline|remove| nodes.
 In the case of \lstinline|add|, the \lstinline|add| node is omited in the true branch of \lstinline|in_rel|, such that packets whose fields are already in the relation will not be seen by the controller due to this \lstinline|add| node.
@@
 \begin{figure}[ht]
-  TODO
+  \includegraphics[scale=.33]{img/optdtree.png}
 \caption{Optimized Decision Tree for Stateful Firewall}     

### 2015-05-10 draft.tex
pretty printing

@@
          pop the top of the stack as pkt
-         run f of pkt resulting in a forwarding decision fd (and a modified packet pkt)
-         record the forwarding decision for the symbolic packet pkt at the current location in the decision tree
+         run f of pkt resulting in a forwarding decision fd and a refined packet pkt
+         record the forwarding decision for the symbolic packet pkt in the decision tree
          set the current location to be the root of the decision tree and repeat
@@
       for each packets in input,
-          find the first of the current rule matching the packet, and record the forwarding decision fd
+          find the first rule matching the packet, and record the forwarding decision fd
           if packet is sent to the controller then
-              evaluate the policy on the packet, potentially adding or removing tuples from relations, record forwarding decision fd'
+              evaluate the policy on the packet, potentially updating relations
               regenerate concrete rules
-              forward the packet according the forwarding decision fd'
+              forward the packet according the evaluated forwarding decision
           else
@@
 % probably no evaluation :(
-\section*{Evaluation}
+% \section*{Evaluation}
 % 1) optimality w.r.t. packets seen by the controler?

### 2015-05-10 draft.tex
..

@@
 
-From an algorithmic policy such as the one included in Fig~\ref{fig:firewall-eg}, we generate a decision tree summarizing the function's interaction with the packet and the relevant decisions that will be made for particular classes of packets.
+From an algorithmic policy such as the one included in Fig~\ref{fig:firewallcode}, we generate a decision tree summarizing the function's interaction with the packet and the relevant decisions that will be made for particular classes of packets.
 The decision tree is created by analyzing the programmers function statically and then used by our runtime system to install switch rules and update them only when necessary (i.e., in response to changes in global state).
@@
 
- For example, after having seen a packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
+  For example, a packing coming from \lstinline|ipsrc| 150 and \lstinline|ipdst| 10 would be matched by the second rule, and dropped on the switch.
+Subsequently, on a packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
 
@@
 \end{lstlisting}
+
+In this example, packet traveling from \lstinline|150| to \lstinline|10| would be matched by the second rule on the switch and delivered, while packets from \lstinline|150| to other internal hosts or from other external host to any internal hosts would be dropped.
 

### 2015-05-10 draft.tex
rule generation

@@
 
+\subsection*{Generating the Symbolic and Concrete Rules}
+
+When \lstinline|compile| terminates, a complete decision tree has been traced for the input algoritmic policy.
+The next step is generate symbolic rules, mapping a symbolic packet to a forwarding decision.
+To do so, we do a depth-first, left to right traversal of the tree.
+On reaching an \lstinline|Add| or a \lstinline|Remove| node, we create a new rule sending the relevant packets to the controller, pruning the rest of the sub-tree.
+On reaching a leaf, we create a new rule forwarding relevant packets according to the forwarding decision held in the leaf node.
+
+We then generate concrete rules from a collection of relations and a list of symbolic rules.
+Concrete rules differ from symbolic rule in that they do not depend on relations.
+Each symbolic rule expands into a list of concrete rules.
+For each of the tuple in the relation on which the rule depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple.
+We repeat this operation on the created list of symbolic rules until all dependant relations are taken care off, at which point the rules are considered concrete.
+Due to the left-to-right order of collection, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.
+
+For example, taking the following symbolic rules: 
+  \begin{lstlisting}
+ipsrc in [101, 1000], (ipsrc,ipdst) in conn
+                  --> Deliver
+ipsrc in [101, 1000], (ipsrc,ipdst) not in conn
+                  --> Drop
+  \end{lstlisting}
+  and the relation ``conn'' as \lstinline| {(150,10), (200, 9), (8, 150)}|, we start by expanding the first rule to:
+\begin{lstlisting}[mathescape]
+ipsrc in ([101, 1000] $\cap$ 150), ipdst = 10
+    --> Deliver
+ipsrc in ([101, 1000] $\cap$ 200), ipdst = 9
+    --> Deliver
+ipsrc in ([101, 1000] $\cap$ 8), ipdst = 150
+    --> Deliver
+ \end{lstlisting}
+which reduces to:
+\begin{lstlisting}
+ipsrc = 150, ipdst = 10
+    --> Deliver
+ipsrc = 200, ipdst = 9
+    --> Deliver
+ \end{lstlisting}
+Note the omission of the third concrete rule, resulting from \lstinline|ipsrc| having an empty range.
+The second symbolic rule is then taken directly as concrete rule, with the packet in the relation having been matched by the previously generated concrete rule, resulting in:
+\begin{lstlisting}
+ipsrc = 150, ipdst = 10
+    --> Deliver
+ipsrc = 200, ipdst = 9
+    --> Deliver  
+ipsrc in [101, 1000]
+    --> Drop
+\end{lstlisting}
+
 \subsection*{Simulating the Network Controller}

### 2015-05-11 draft.tex
..
reformatted code and examples

@@
 
-\title{\textit{Red Oak}: Proactive Policy Generation}
+\title{\textit{Red Oak}\\ Proactive Controller Generation for Algorithmic Policy}
  \authorinfo{Ryan Beckett}{Princeton University}{rbeckett@princeton.edu}
@@
 In particular, for each IP source that is trusted to communicate with an internal destination in the "conn" relation, we will get a concrete rule permitting this communication.
-  \begin{lstlisting}
-ipsrc in [0, 100] --> Sent to Controller
-ipsrc in [101, 1000], (ipsrc,ipdst) in conn
-                  --> Deliver
-ipsrc in [101, 1000], (ipsrc,ipdst) not in conn
-                  --> Drop
+  \begin{lstlisting}[mathescape]
+src $\in$ [0, 100] $\mapsto$ Controller
+src $\in$ [101, 1000], (src,dst) $\in$ conn  $\mapsto$ Deliver
+src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
   \end{lstlisting}
@@
 After initialization (starting with no trusted connections), the concrete rules would be generated from the symbolic rules by having all relations be empty:
-\begin{lstlisting}
-ipsrc in [0, 100] ---> Sent to Controller  
-ipsrc in [101, 1000] ---> Drop  
+\begin{lstlisting}[mathescape]
+src $\in$ [0, 100] $\mapsto$ Controller  
+src $\in$ [101, 1000] $\mapsto$ Drop  
 \end{lstlisting}
@@
 
-  For example, a packing coming from \lstinline|ipsrc| 150 and \lstinline|ipdst| 10 would be matched by the second rule, and dropped on the switch.
-Subsequently, on a packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
+  For example, a packing coming from \lstinline|src| 150 and \lstinline|dst| 10 would be matched by the second rule, and dropped on the switch.
+Subsequently, on a packet with \lstinline|src| 10 and \lstinline|dst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
 
-\begin{lstlisting}
-ipsrc in [0, 100] ---> Sent to Controller  
-ipsrc = 150, ipdst = 10 ---> Deliver  
-ipsrc in [101, 1000] ---> Drop  
+\begin{lstlisting}[mathescape]
+src $\in$ [0, 100] $\mapsto$ Controller  
+src = 150, dst = 10 $\mapsto$ Deliver  
+src $\in$ [101, 1000] $\mapsto$ Drop  
 \end{lstlisting}
@@
 \begin{lstlisting}
-let f_cross (f_1) (f_2) pkt =
-  let d_1 = f_1 pkt in
-  let d_2 = f_2 pkt in
-  Multicast (d_1,d_2)
+let f_cross (f1) (f2) pkt =
+  let d1 = f1 pkt in
+  let d2 = f2 pkt in
+  Multicast (d1,d2)
 \end{lstlisting}
@@
 \begin{figure}[ht]
-  \begin{lstlisting}
-type packet
+  \begin{lstlisting}[mathescape]
+type pkt
+type field = IpSrc | IpDst 
+type fields = field list
+type range = (int*int) list
 
-type field = IpSrc | IpDst
-
-type range = (int*int) list
+type decision =
+	| Deliver
+	| Drop
+	| ForwardTo of int
+	| Multicast of decision * decision
 
@@
 
-type forwarding_decision =
-	| Deliver
-	| Drop
-	| ForwardTo of int
-	| Multicast of forwarding_decision
-                      * forwarding_decision
-
-val in_range: packet -> field -> range -> bool
+val in_range: pkt $\rightarrow$ field $\rightarrow$ range $\rightarrow$ bool
 \end{lstlisting}
@@
 \begin{figure}[ht]
-  \begin{lstlisting}
+  \begin{lstlisting}[mathescape]
 type relation = string
 
-val add: packet -> field list -> relation
-    -> unit
-
-val remove: packet -> field list -> relation
-    -> unit
-
-val in_rel: packet -> field list -> relation
-    -> bool
+val add: pkt $\rightarrow$ fields $\rightarrow$ relation $\rightarrow$ unit
+val remove: pkt $\rightarrow$ fields $\rightarrow$ relation $\rightarrow$ unit
+val in_rel: pkt $\rightarrow$ fields $\rightarrow$ relation $\rightarrow$ bool
 \end{lstlisting}
@@
 \begin{figure}[ht]
-  \begin{lstlisting}
-val compile: (packet -> forwarding_decision)
-    -> policy
+  \begin{lstlisting}[mathescape]
+type examples =  ((field * int) list) list
 
-val run: policy -> ((field * int) list) list
-    -> unit
-
-val print_policy: policy -> unit
+val compile: (pkt $\rightarrow$ decision) $\rightarrow$ policy
+val run: policy $\rightarrow$ examples $\rightarrow$ unit
   \end{lstlisting}
@@
 For example, taking the following symbolic rules: 
-  \begin{lstlisting}
-ipsrc in [101, 1000], (ipsrc,ipdst) in conn
-                  --> Deliver
-ipsrc in [101, 1000], (ipsrc,ipdst) not in conn
-                  --> Drop
+  \begin{lstlisting}[mathescape]
+src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
+src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
   \end{lstlisting}
@@
 \begin{lstlisting}[mathescape]
-ipsrc in ([101, 1000] $\cap$ 150), ipdst = 10
-    --> Deliver
-ipsrc in ([101, 1000] $\cap$ 200), ipdst = 9
-    --> Deliver
-ipsrc in ([101, 1000] $\cap$ 8), ipdst = 150
-    --> Deliver
+src $\in$ ([101, 1000] $\cap$ 150), dst = 10 $\mapsto$ Deliver
+src $\in$ ([101, 1000] $\cap$ 200), dst = 9 $\mapsto$ Deliver
+src $\in$ ([101, 1000] $\cap$ 8), dst = 150 $\mapsto$ Deliver
  \end{lstlisting}
 which reduces to:
-\begin{lstlisting}
-ipsrc = 150, ipdst = 10
-    --> Deliver
-ipsrc = 200, ipdst = 9
-    --> Deliver
+\begin{lstlisting}[mathescape]
+src = 150, dst = 10 $\mapsto$ Deliver
+src = 200, dst = 9 $\mapsto$ Deliver
  \end{lstlisting}
@@
 The second symbolic rule is then taken directly as concrete rule, with the packet in the relation having been matched by the previously generated concrete rule, resulting in:
-\begin{lstlisting}
-ipsrc = 150, ipdst = 10
-    --> Deliver
-ipsrc = 200, ipdst = 9
-    --> Deliver  
-ipsrc in [101, 1000]
-    --> Drop
+\begin{lstlisting}[mathescape]
+src = 150, dst = 10 $\mapsto$ Deliver
+src = 200, dst = 9 $\mapsto$ Deliver
+src $\in$ [101, 1000] $\mapsto$ Drop
 \end{lstlisting}
@@
 From which the following symbolic rule would be generate:
-\begin{lstlisting}
-ipsrc in [0, 100], (ipdst, ipsrc) in conn
-                  --> Deliver
-ipsrc in [0, 100] --> Sent to Controller
-ipsrc in [101, 1000], (ipsrc,ipdst) in conn
-                  --> Deliver
-ipsrc in [101, 1000], (ipsrc,ipdst) not in conn
-                  --> Drop
+\begin{lstlisting}[mathescape]
+src $\in$ [0, 100], (dst, src) in conn $\mapsto$ Deliver
+src $\in$ [0, 100] $\mapsto$ Controller
+src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
+src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
 \end{lstlisting}
@@
 However, after seeing, for example, packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the rules would be
-\begin{lstlisting}
-ipsrc = 10, ipdst = 150 ---> Deliver  
-ipsrc in [0, 100] ---> Sent to Controller  
-ipsrc = 150, ipdst = 10 ---> Deliver  
-ipsrc in [101, 1000] ---> Drop  
+\begin{lstlisting}[mathescape]
+src = 10, dst = 150 $\mapsto$ Deliver  
+src $\in$ [0, 100] $\mapsto$ Controller  
+src = 150, dst = 10 $\mapsto$ Deliver  
+src $\in$ [101, 1000] $\mapsto$ Drop  
 \end{lstlisting}

### 2015-05-11 draft.tex
change pseudocode

@@
 \usepackage{tikz-qtree}
+
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \begin{figure*}[ht]
-\begin{lstlisting}
-let compile (f: packet -> forwarding_decision) : policy =
+\begin{lstlisting}[mathescape]
+let compile (f: pkt $\rightarrow$ forwarding_decision) : policy =
       push most general packet to (empty) stack
@@
 \begin{lstlisting}
-let run (pol: policy) (inputs: ((field*int) list) list) : unit =
+let run (pol: policy) (inputs: examples) : unit =
       build the symbolic rules for policy pol

### 2015-05-11 draft.tex
internal representation

@@
 We include in Figure~\ref{fig:range_api} the core of our \lstinline|A.P.I.|.
-\lstinline|field| enumerates the different fields of a packet.
-It can be extended to represent more fields of a packet's header.
-Associated with \lstinline|field| are \lstinline|range|s, possible values taken by a certain packet's field, within a given minimum value \lstinline|min_value| and maximum value \lstinline|max_value|.
-\lstinline|range| consists of list of integer pairs, representing a collection of intervals whose union correspond to the range of values of a packet.
-For simplicity, we consider integer fields here, but this could be extended, for example, to IP prefix, boolean flags, etc.
-
-\lstinline|forwarding_decision| enumerates the different forwarding decision which can be returned by an algorithmic policy.
-\lstinline|Deliver| forwards the packet to its destination, \lstinline|Drop| drops the packet, \lstinline|ForwardTo i| forwards the packet to a constant location \lstinline|i| and \lstinline|Multicast fd1 fd2| duplicates the packet and handles each according to \lstinline|fd1| and \lstinline|fd2| respectively.
-
-Algorithmic policy may base their forwarding decision on the result of calling the \lstinline|in_range pkt fld r|, which returns whether value of packet \lstinline|pkt| at field \lstinline|fld| is in the range \lstinline|r|.
 
@@
 
-In Figure~\ref{fig:rel_api}, we include the relation portion of the \lstinline|A.P.I.|, used to model stateful controller programs.
-A \lstinline|relation| represents a table in which each row contains a series of field values.
-It is referred to by a string.
-Policy writer may add a row to a relation \lstinline|rel| by calling the function \lstinline|add pkt flds rel|, which will populate the new row with the fields \lstinline|flds|' values of the current packet \lstinline|pkt|.
-Similarly, they may remove a row containing exactly the values of fields \lstinline|flds|' values of the current packet \lstinline|pkt| by calling \lstinline|remove pkt flds rel|.
-Finally, one may test if the values of fields \lstinline|flds| of current packet \lstinline|pkt| are in relation \lstinline|rel| by calling \lstinline|in_rel pkt flds rel|.
+\lstinline|field| enumerates the different fields of a packet.
+It can be extended to represent more fields of a packet's header.
+Associated with \lstinline|field| are \lstinline|range|s, possible values taken by a certain packet's field, within a given minimum value \lstinline|min_value| and maximum value \lstinline|max_value|.
+\lstinline|range| consists of list of integer pairs, representing a collection of intervals whose union correspond to the range of values of a packet.
+For simplicity, we consider integer fields here, but this could be extended, for example, to IP prefix, boolean flags, etc.
+
+\lstinline|forwarding_decision| enumerates the different forwarding decision which can be returned by an algorithmic policy.
+\lstinline|Deliver| forwards the packet to its destination, \lstinline|Drop| drops the packet, \lstinline|ForwardTo i| forwards the packet to a constant location \lstinline|i| and \lstinline|Multicast fd1 fd2| duplicates the packet and handles each according to \lstinline|fd1| and \lstinline|fd2| respectively.
+
+Algorithmic policy may base their forwarding decision on the result of calling the \lstinline|in_range pkt fld r|, which returns whether value of packet \lstinline|pkt| at field \lstinline|fld| is in the range \lstinline|r|.
 
@@
 
+In Figure~\ref{fig:rel_api}, we include the relation portion of the \lstinline|A.P.I.|, used to model stateful controller programs.
+A \lstinline|relation| represents a table in which each row contains a series of field values.
+It is referred to by a string.
+Policy writer may add a row to a relation \lstinline|rel| by calling the function \lstinline|add pkt flds rel|, which will populate the new row with the fields \lstinline|flds|' values of the current packet \lstinline|pkt|.
+Similarly, they may remove a row containing exactly the values of fields \lstinline|flds|' values of the current packet \lstinline|pkt| by calling \lstinline|remove pkt flds rel|.
+Finally, one may test if the values of fields \lstinline|flds| of current packet \lstinline|pkt| are in relation \lstinline|rel| by calling \lstinline|in_rel pkt flds rel|.
+
 \begin{figure}[ht]
@@
 
+The rest of the API, included in Figure~\ref{fig:build_api}, consists of functions used to interact with Red Oak.
+\lstinline|compile| is used to trace a policy from a given function from packet to forwarding \lstinline|decision|.
+\lstinline|run| simulates the compiled \lstinline|policy| on a series of \lstinline|examples|.
+In the remainder of this section, we describe these functions in more detail.
+
+\subsection*{Internal Representation of Packets and Policies}
+
+Packet are represented abstractly in the API as type \lstinline|pkt|, and are treated as concrete packet by algorithmic policies.
+In contrast, internally, \lstinline|pkt| represents a class of packet which could reach a certain point in the program.
+We call this class of packet a symbolic packet.
+A symbolic packet consists of two parts:
+\begin{itemize}
+\item a map from \lstinline|field| to \lstinline|range|, representing the range of possible value from the class of packet
+  \item a list of pairs of \lstinline|relation| and list of \lstinline|field| representing the relation that the relevant \lstinline|field|'s value of the symbolic packet should be in. 
+  \end{itemize}
+
+A \lstinline|policy| is represented internally as a decision tree whose leaves are forwarding decision and whose node are either branching according the range of a field value, according to a relation, or adding and removing components from a relation.
+To simplify rule generation, we also record in the tree the most general symbolic packet who can reach leaves, \lstinline|Add| and \lstinline|Remove| nodes.
+
 \subsection*{Tracing the Algorithmic Policy}
-From an algorithmic policy, we generate a decision tree  whose leaves are forwarding decision and whose node are either branching according the range of a field value, according to a relation ,or adding and removing components from a relation.
+From an algorithmic policy, we generate a decision tree using the function \lstinline|compile|.
 

### 2015-05-11 draft.tex
beginning abstract
abstract

@@
 
-\begin{abstract} ...
+\begin{abstract}
+Software Defined Networking (SDN) is a new paradigm that gives programmers 
+centralized visibility and control over the network. However, due to performance 
+demands, writing correct and efficient programs is challenging. Programmers must 
+reason about how to correctly forward packets in the network, how to maintain
+global network state, which packets to observe at the centralized controller, 
+and how to minimize latency by pushing as much decision logic
+as possible onto the high-performance switch elements.
+
+We introduce a high-level SDN library that can be embedded into existing languages.
+Our library exposes an API that allows programmers to directly express behavioral intent, 
+while leaving the performance details to the library's optimization algorithms.
+Our library supports basic tests on packet fields as well as a simple mutable table
+data structure for storing global network information (e.g., firewall ACL, MAC learning table).
+Finally, to ensure adequate performance, 
+we statically analyze the programmer's policy by pushing symbolic packets through the policy 
+and observing the resulting packet forwarding decision.
+
 \end{abstract}

### 2015-05-11 draft.tex
abstract
conclusion

@@
 and observing the resulting packet forwarding decision.
-
 \end{abstract}
@@
 \section*{Future Work and Conclusion}
+In this paper, we have introduced Oak, a SDN library for proactively switch forwarding rules
+from an arbitrary programmer-defined policy in a general purpose programming language.
+Oak provides programmers the abstraction that their centralized policy will run on each 
+new packet entering the network, while alleviating for the programmer need to consider
+how to split the decision logic between to switches and the controller.
+
+The Oak API exposes simple tests on packet fields as well as a global, mutable table data
+structure for recording information about previous packets observed in the network.
+To ensure that most packets are handled directly on the switches, we employ a static analysis
+that pushes symbolic packets through the programmer's policy to explore how all classes of 
+packets will be handled, determining which packets must be observed by the controller and which 
+packets need not be. Finally, we describe our Oak prototype's run-time system that 
+specializes decisions after making concrete observations in the network that dictate forwarding policy.
 

### 2015-05-11 draft.tex
..

--- 2015-05-11 draft.tex
@@
 \begin{abstract}
-Software Defined Networking (SDN) is a new paradigm that gives programmers 
+Software Defined Networking (SDN) is a paradigm that gives programmers 
 centralized visibility and control over the network. However, due to performance 
@@
 \section*{Future Work and Conclusion}
+In this paper, we have introduced \textit{Red Oak}, a SDN library for proactively generating a dynamic controller from an arbitrary programmer-defined policy in a general purpose programming language.
+\textit{Red Oak} provides programmers the abstraction that their centralized policy will run on each 
+new packet entering the network, while alleviating for the programmer need to consider
+how to split the decision logic between to switches and the controller.
+
+The \textit{Red Oak} API exposes simple tests on packet fields as well as a global, mutable table data
+structure for recording information about previous packets observed in the network.
+To ensure that most packets are handled directly on the switches, we employ a static analysis
+that pushes symbolic packets through the programmer's policy to explore how all classes of 
+packets will be handled, determining which packets must be observed by the controller and which 
+packets need not be. Finally, we describe our \textit{Red Oak} prototype's run-time system that 
+specializes decisions after making concrete observations in the network that dictate forwarding policy.
 

### 2015-05-11 draft.tex
add comment on pre populated relations

@@
 
+While we currently start the simulation with empty relation, \lstinline|run| could easily be extended to initialize the controller with pre-populated relations.
+
 \subsection*{Optimizations}
@@
 
-\section*{Future Work and Conclusion}
+\section*{Conclusion}
 In this paper, we have introduced \textit{Red Oak}, a SDN library for proactively generating a dynamic controller from an arbitrary programmer-defined policy in a general purpose programming language.

### 2015-05-11 draft.tex
..

@@
 
-We introduce a high-level SDN library that can be embedded into existing languages.
+We developed a high-level SDN library that can be embedded into existing languages.
 Our library exposes an API that allows programmers to directly express behavioral intent, 
@@
 The typical model for SDN involves a centralized controller or distributed set of controllers programming switches dynamically to implement the desired control functionality.
-However, the questions remains as to how best to write programs for SDNs.
+However, the question remains as to how best to write programs for SDNs.
 
@@
 
-Ultimately, the performance requirements of networks dictate that the centralized controller(s) can not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency compared to those being handled by switches.
+Ultimately, the performance requirements of networks dictate that the centralized controller(s) should not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency compared to those being handled by switches.
 For this reason, compiler technologies for network programming languages must be able to push much of the control plane functionality into the data plane switches themselves.
 
-For Maple, the authors show how to optimize the controller so that packets are handled without having to run the function on them.
+In Maple \cite{Maple}, the authors show how to optimize the controller to reduce the number of packets are handled by running the function on them.
 Maple does so by providing a packet API capturing each access to packet fields, and memoizing the results of the function for subsequent packets sharing the accessed fields' values.

### 2015-05-11 draft.tex
typo

@@
 \begin{enumerate}
-\item Since controller decisions are memoized, all initial packets must first go to the controller before it can learn how to forward similar packets directly in the data plane.
+\item Since controller decisions are memoized at runtime, all packets must initially go to the controller before it can learn how to forward similar packets directly in the data plane.
 This leads to a cold-startup phase with high packet latency.
@@
 
-In this work, we language-independent API and compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
-However, we address each of the shortcomings in Maple listed above. Our contributions are the following:
+In this work, we devised a language-independent API together with a compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
+However, we address each of the shortcomings  listed above. Our contributions are the following:
 
 \begin{enumerate}
-\item We describe a static analysis on the user-defined forwarding policy that pushes symbolic packets through the function to explore all possible paths of execution.
-This analysis is then used to proactively generate forwarding rules to install in the network.
-\item We provide a packet API that lets programmers ask if a field is in a particular range (e.g., IP prefix) rather than if the the field equals a particular value.
+\item We describe a static analysis on the user-defined forwarding policy that explore all possible paths of execution by evaluating the function on symbolic packets.
+This analysis is then used to proactively generate a network controller and initial forwarding rules to install in the network.
+\item We provide a packet API that lets programmers ask if a field is in a particular range (e.g.
+an IP prefix) rather than if the field equals a particular value.
 This more expressive test works together with our analysis to better capture programmer intent when setting up forwarding decision in the data plane.

### 2015-05-11 draft.tex
..

@@
    % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to dst)
-	The programming model for Oak is that each packet entering the network is mapped to a forwarding decision that describes where the packet should end up leaving the network.
+	The programming model for \textit{Red Oak} is that each packet entering the network is mapped to a forwarding decision that describes where the packet should end up leaving the network.
 For simplicity, we assume that we are targeting a one big switch implementation of the network.
@@
 
-The rest of the API, included in Figure~\ref{fig:build_api}, consists of functions used to interact with Red Oak.
+The rest of the API, included in Figure~\ref{fig:build_api}, consists of functions used to interact with \textit{Red Oak}.
 \lstinline|compile| is used to trace a policy from a given function from packet to forwarding \lstinline|decision|.

### 2015-05-11 draft.tex
edits first pass

--- 2015-05-11 draft.tex
@@
 
-We developed a high-level SDN library that can be embedded into existing languages.
+We introduce a high-level SDN library that can be embedded into existing general-purpose languages.
 Our library exposes an API that allows programmers to directly express behavioral intent, 
@@
 The typical model for SDN involves a centralized controller or distributed set of controllers programming switches dynamically to implement the desired control functionality.
-However, the question remains as to how best to write programs for SDNs.
+However, the question remains as how best to write programs for SDNs.
 
@@
 
-Ultimately, the performance requirements of networks dictate that the centralized controller(s) should not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency compared to those being handled by switches.
+Ultimately, the performance requirements of networks dictate that the centralized controller(s) can not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency compared to those being handled by switches.
 For this reason, compiler technologies for network programming languages must be able to push much of the control plane functionality into the data plane switches themselves.
 
-In Maple \cite{Maple}, the authors show how to optimize the controller to reduce the number of packets are handled by running the function on them.
+For Maple, the authors show how to optimize the controller so that packets are handled without having to run the function on them.
 Maple does so by providing a packet API capturing each access to packet fields, and memoizing the results of the function for subsequent packets sharing the accessed fields' values.
@@
 
-In this work, we language-independent API and compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
+In this work, we develop a language-independent API and compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
 However, we address each of the shortcomings in Maple listed above. Our contributions are the following:
@@
 This analysis is then used to proactively generate forwarding rules to install in the network.
-\item We provide a packet API that lets programmers ask if a field is in a particular range (e.g., IP prefix) rather than if the the field equals a particular value.
-This more expressive test works together with our analysis to better capture programmer intent when setting up forwarding decision in the data plane.
+\item We provide a packet API that lets programmers ask if a field is in a particular range (e.g., IP prefix) rather than if the field equals a particular value.
+This more expressive test works together with our analysis to better capture programmer intent when setting up forwarding decisions in the data plane.
 \item We provide a mutable relation data structure in our API to maintain global state.
@@
 Conceptually this model offers the abstraction that the network forwards packets as if this function were executed on each packet when entering the network.
-To handle future traffic directly in the data plane, Maple keeps a tree data structure that branches on decisions observed at run time (e.g., the packet had source IP 10.0.0.1)
+To handle future traffic directly in the data plane, Maple keeps a tree data structure that branches on decisions observed at run time (e.g., the packet had source IP 10.0.0.1).
+We extend the idea of decision tree to work at a higher level of abstraction involving decisions based on symbolic constraints, rather than concrete, packets.
+Let us take a look at an example.
 
 A policy for a simple stateful firewall using our API is included in Figure~\ref{fig:firewallcode}.
-For simplicity, we assume that IP ranges are between 0 and 1000, however it is easy to generalize this to actual IP prefixes.
+For simplicity, we assume that IP addresses range between 0 and 1000, however it is easy to generalize this to actual IP prefixes.
 In the example, source IPs from 0 to 10 are considered to be internal to the network and everything else, external.
 
-If a packet is sent from an internal host to an external host, the pair of internal and external host is added to a list of trusted connection.
+If a packet is sent from an internal host to an external host, the pair of internal and external host is added to a list of trusted connections.
 If a packet is received from an external host, we lookup in the table to see if the connection is trusted, in which case we forward the packet to its destination, otherwise we drop it.
@@
 From an algorithmic policy such as the one included in Fig~\ref{fig:firewallcode}, we generate a decision tree summarizing the function's interaction with the packet and the relevant decisions that will be made for particular classes of packets.
-The decision tree is created by analyzing the programmers function statically and then used by our runtime system to install switch rules and update them only when necessary (i.e., in response to changes in global state).
+The decision tree is created by analyzing the programmers function statically and then used by our run time system to install switch rules and update them only when necessary (i.e., in response to changes in global state).
 For the stateful firewall included in figure~\ref{fig:firewallcode}, the following decision tree would be generated:
@@
   For example, a packing coming from \lstinline|src| 150 and \lstinline|dst| 10 would be matched by the second rule, and dropped on the switch.
-Subsequently, on a packet with \lstinline|src| 10 and \lstinline|dst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:
+Subsequently, on a packet with \lstinline|src| 10 and \lstinline|dst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would be updated as follows:
 
@@
 
-In this example, packet traveling from \lstinline|150| to \lstinline|10| would be matched by the second rule on the switch and delivered, while packets from \lstinline|150| to other internal hosts or from other external host to any internal hosts would be dropped.
+In this example, a packet traveling from \lstinline|150| to \lstinline|10| would be matched by the second rule on the switch and delivered, while packets from \lstinline|150| to other internal hosts or from other external host to any internal hosts would be dropped.
 
@@
 For simplicity, we assume that we are targeting a one big switch implementation of the network.
-The one big switch model has been heavily optimized \cite{Obs}, and we view the problem of determining paths through the network as orthogonal to our work here.
+The one big switch model has been heavily optimized (e.g., \cite{Obs}), and we view the problem of determining paths through the network as orthogonal to our work here.
 
@@
 The programmer can deliver a packet, according to the packet's desination, drop a packet, and forwarding a packet to a static location.
-Our model also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller (not shown in Fig~\ref{fig:range_api}).
+Our API also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller (not shown in Fig~\ref{fig:range_api}).
 Finally, the programmer can combine several forwarding decisions into one by using a mulitcast forwarding decision.
@@
 
-   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_src_range| a partial policy, which, given a policy \lstinline|f| and a range \lstinline|r|, will forwards all packets whose source IP is within range \lstinline|r| and drop the packet otherwise.
+   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_src_range| a partial policy, which, given a policy \lstinline|f| and a range \lstinline|r|, will forward all packets whose source IP is within range \lstinline|r| and drop the packet otherwise.
 The ability to compose two or more policies in sequence like this arises many times in practice.
@@
 
-Various composition operator may be built as host language function to combine multiple algorithmic policies into a single one.
+Various composition operator may be built as host language functions to combine multiple algorithmic policies into a single one.
 For example, \lstinline|f_cross|, included in Figure~\ref{fig:ex-cross}, runs algorithmic policy \lstinline|f_1| and \lstinline|f_2| on input packets and multicasts them according to both forwarding decisions.
@@
 \subsection*{A.P.I.}
-In our model, algorithmic policies interact with packets through A.P.I function calls before returning a forwarding decision.
-We include in Figure~\ref{fig:range_api} the core of our \lstinline|A.P.I.|.
+In our programming model, algorithmic policies interact with packets through API function calls before returning a forwarding decision.
+We include in Figure~\ref{fig:range_api} the core of our \lstinline|API.|.
 
@@
 
-\lstinline|forwarding_decision| enumerates the different forwarding decision which can be returned by an algorithmic policy.
+\lstinline|decision| enumerates the different forwarding decision which can be returned by an algorithmic policy.
 \lstinline|Deliver| forwards the packet to its destination, \lstinline|Drop| drops the packet, \lstinline|ForwardTo i| forwards the packet to a constant location \lstinline|i| and \lstinline|Multicast fd1 fd2| duplicates the packet and handles each according to \lstinline|fd1| and \lstinline|fd2| respectively.
 
-Algorithmic policy may base their forwarding decision on the result of calling the \lstinline|in_range pkt fld r|, which returns whether value of packet \lstinline|pkt| at field \lstinline|fld| is in the range \lstinline|r|.
+Algorithmic policies may base their forwarding decision on the result of calling the \lstinline|in_range pkt fld r|, which returns whether value of packet \lstinline|pkt| at field \lstinline|fld| is in the range \lstinline|r|.
 
@@
 
-In Figure~\ref{fig:rel_api}, we include the relation portion of the \lstinline|A.P.I.|, used to model stateful controller programs.
+In Figure~\ref{fig:rel_api}, we include the relation portion of the \lstinline|API|, used to model stateful controller programs.
 A \lstinline|relation| represents a table in which each row contains a series of field values.
 It is referred to by a string.
-Policy writer may add a row to a relation \lstinline|rel| by calling the function \lstinline|add pkt flds rel|, which will populate the new row with the fields \lstinline|flds|' values of the current packet \lstinline|pkt|.
+The policy writer may add a row to a relation \lstinline|rel| by calling the function \lstinline|add pkt flds rel|, which will populate the new row with the fields \lstinline|flds|' values of the current packet \lstinline|pkt|.
 Similarly, they may remove a row containing exactly the values of fields \lstinline|flds|' values of the current packet \lstinline|pkt| by calling \lstinline|remove pkt flds rel|.
@@
 The rest of the API, included in Figure~\ref{fig:build_api}, consists of functions used to interact with Red Oak.
-\lstinline|compile| is used to trace a policy from a given function from packet to forwarding \lstinline|decision|.
-\lstinline|run| simulates the compiled \lstinline|policy| on a series of \lstinline|examples|.
+The \lstinline|compile| function is used to symbolically analyze a user-function for forwarding packets and generate, from this function, a decision tree that is captured as an abstract type called policy.
+The \lstinline|run| function simulates the compiled \lstinline|policy| on a series of example packets entering the network.
+This incorporates the Red Oak run time system to update forwarding decisions based on the global state.
 In the remainder of this section, we describe these functions in more detail.
@@
 
-Packet are represented abstractly in the API as type \lstinline|pkt|, and are treated as concrete packet by algorithmic policies.
-In contrast, internally, \lstinline|pkt| represents a class of packet which could reach a certain point in the program.
-We call this class of packet a symbolic packet.
+Packets are represented abstractly in the API as type \lstinline|pkt|, and are treated as concrete packets by algorithmic policies.
+In contrast, internally, \lstinline|pkt| represents a class of packets which could reach a certain point in the program.
+We call this class of packets a symbolic packet.
 A symbolic packet consists of two parts:
 \begin{itemize}
-\item a map from \lstinline|field| to \lstinline|range|, representing the range of possible value from the class of packet
+\item a map from \lstinline|field| to \lstinline|range|, representing the range of possible value for the class of packets
   \item a list of pairs of \lstinline|relation| and list of \lstinline|field| representing the relation that the relevant \lstinline|field|'s value of the symbolic packet should be in. 
@@
 
-A \lstinline|policy| is represented internally as a decision tree whose leaves are forwarding decision and whose node are either branching according the range of a field value, according to a relation, or adding and removing components from a relation.
-To simplify rule generation, we also record in the tree the most general symbolic packet who can reach leaves, \lstinline|Add| and \lstinline|Remove| nodes.
+A \lstinline|policy| is represented internally as a decision tree whose leaves are forwarding decision and whose nodes are either branching according to the range of a field value, branching according to a membership in a relation, or adding and removing components from a relation.
+To simplify rule generation, we also record, at each node, the most general symbolic packet that can reach that node.
+The intuition is that any packets that can reach an \lstinline|add| or \lstinline|remove| node, must be observed by the controller.
 
 \subsection*{Tracing the Algorithmic Policy}
-From an algorithmic policy, we generate a decision tree using the function \lstinline|compile|.
+From an algorithmic policy, we generate a decision tree using the function \lstinline|compile|. The pseudo code is shown in Figure~\ref{fig:compile-pseudo}
 
@@
 
-When \lstinline|f| calls \lstinline|in_range| to learn if the field \lstinline|fld|'s value is in the range \lstinline|r|, we
+When \lstinline|f| calls \lstinline|in_range| to learn if the packet \lstinline|fld|'s value is in the range \lstinline|r|, we
 \begin{enumerate}
-\item  create an \lstinline|Inrange| branching node in the decision tree
-\item refine the current symbolic packet to represent all packets in range \lstinline|r| that could reach this test in the algorithmic policy and return true
-\item push on the stack of packets to process a symbolic packets representing all packets in the complement of \lstinline|r| for field \lstinline|fld|.
+\item  Create a new \lstinline|in_range| branching node in the decision tree with subtrees corresponding to \lstinline|true| and \lstinline|false|
+\item Refine the current symbolic packet further to intersect the tested range. This corresponds to exploring the \lstinline|true| branch of the decision tree. We return true. 
+\item Push onto a stack consisting of remaining packets to process, a new symbolic packet that refines the range of the original symbolic packet to explore the false branch of the current range test by intersecting the complement of \lstinline|r| for field \lstinline|fld| with the current symbolic packet.
 \end{enumerate}
@@
 \begin{enumerate}
-  \item create an \lstinline|Inrelation| branching node in the decision tree
-  \item tag the current packet as being in the relation, return true
-  \item push on the stack of packets to process the same packet tagged as not being part of relation \lstinline|rel|.
+  \item Create a new \lstinline|in_rel| branching node in the decision tree wit hsubtrees corresponding to \lstinline|true| and \lstinline|false|
+  \item Tag the current packet with the symbolic constraint of being in the relation and return true
+  \item Push onto a stack consisting of remaining packets to process, a new symbolic that is the same as the current packet, but tagged as not being part of the relation \lstinline|rel|.
 \end{enumerate}
@@
 
-While we currently start the simulation with empty relation, \lstinline|run| could easily be extended to initialize the controller with pre-populated relations.
-
 \subsection*{Optimizations}
@@
 
-\section*{Conclusion}
-In this paper, we have introduced \textit{Red Oak}, a SDN library for proactively generating a dynamic controller from an arbitrary programmer-defined policy in a general purpose programming language.
-\textit{Red Oak} provides programmers the abstraction that their centralized policy will run on each 
+\section*{Future Work and Conclusion}
+In this paper, we have introduced Oak, a SDN library for proactively switch forwarding rules
+from an arbitrary programmer-defined policy in a general purpose programming language.
+Oak provides programmers the abstraction that their centralized policy will run on each 
 new packet entering the network, while alleviating for the programmer need to consider
@@
 
-The \textit{Red Oak} API exposes simple tests on packet fields as well as a global, mutable table data
+The Oak API exposes simple tests on packet fields as well as a global, mutable table data
 structure for recording information about previous packets observed in the network.
@@
 packets will be handled, determining which packets must be observed by the controller and which 
-packets need not be. Finally, we describe our \textit{Red Oak} prototype's run-time system that 
+packets need not be. Finally, we describe our Oak prototype's run-time system that 
 specializes decisions after making concrete observations in the network that dictate forwarding policy.

### 2015-05-11 draft.tex
edits 2nd pass

@@
 \section*{Implementation}
-%that's the hope...
+
+We now describe our prototype implementation of \textit{Red Oak} as a library embedded in the Ocaml programming language.
 
@@
 We then generate concrete rules from a collection of relations and a list of symbolic rules.
-Concrete rules differ from symbolic rule in that they do not depend on relations.
+Concrete rules differ from symbolic rules in that they do not depend on relations.
 Each symbolic rule expands into a list of concrete rules.
-For each of the tuple in the relation on which the rule depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple.
-We repeat this operation on the created list of symbolic rules until all dependant relations are taken care off, at which point the rules are considered concrete.
-Due to the left-to-right order of collection, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.
+For each tuple in the relation on which the symbolic rules depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple.
+We repeat this operation on the created list of symbolic rules until all dependant relations are taken care of, at which point the rules are considered concrete.
+Due to the prioritization of the resulting rules, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.
 
-For example, taking the following symbolic rules: 
+For example, taking the following symbolic rules from our earlier stateful firewall example: 
   \begin{lstlisting}[mathescape]
@@
   \end{lstlisting}
-  and the relation ``conn'' as \lstinline| {(150,10), (200, 9), (8, 150)}|, we start by expanding the first rule to:
+The first rules says: "external hosts that are trusted should be let through", and the second rules says: "traffic from untrusted external hosts should be dropped".
+
+  Assume that the trusted connection relation ``conn'' contains the pairs 
+  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external src 150 can talk to internal dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
 \begin{lstlisting}[mathescape]
@@
  \end{lstlisting}
-which reduces to:
+To satisfy both symbolic constraints, we take the intersection of the constraints, which corresponds to intersection of the constraints on the ranges of each field. These rules reduce to:
 \begin{lstlisting}[mathescape]
@@
 Note the omission of the third concrete rule, resulting from \lstinline|ipsrc| having an empty range.
-The second symbolic rule is then taken directly as concrete rule, with the packet in the relation having been matched by the previously generated concrete rule, resulting in:
+We can now concretize the second symbolic rule without doing any work, since checking for relation membership is already done by all the higher priority rules.
+The final concrete forwarding rules are now:
 \begin{lstlisting}[mathescape]
@@
 \lstinline|run| starts by building the symbolic rules for the given policy, and generating the concrete rules assuming all relations to be empty.
-In sequence, packet are matched with the current concrete rules, recording the forwarding decision.
+In sequence, packets are matched with the current concrete rules, recording the forwarding decision.
 If a packet is sent to the controller, we evaluate the policy on the packet, potentially adding or removing tuples from relations.
@@
 For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|ipsrc| of 10 and \lstinline|ipdst| of 150, any subsequent packet with the same \lstinline|ipsrc| and \lstinline|ipdst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent.
-By using both local and global transformations on the decision tree, we can get closer to an optimal amount of packet forwarded to the controller, where only packets modifying relations are seen by the controller.
+By using both local and global transformations on the decision tree, we can get closer to an optimal number of packet forwarded to the controller, where only packets having a real, meaningful effect are seen by the controller.
 

### 2015-05-11 draft.tex
final edits

@@
 In our programming model, algorithmic policies interact with packets through API function calls before returning a forwarding decision.
-We include in Figure~\ref{fig:range_api} the core of our \lstinline|API.|.
+We include in Figure~\ref{fig:range_api} the core of our \lstinline|API|.
 
@@
 \section*{Related Work}
-% deeper comparison with Maple
+There are a number of existing programming abstractions for SDN.
+The Frenetic~\citep{Frenetic} family of languages provide domain-specific operators for composing packet processing functions together and compiling them to switches.
+Here we focus on compiling arbitrary programmer functions in a general-purpose language.
+Merlin~\cite{Merlin} is a language for configuring networks based on path and quality of service constraints.
+However, Merlin focuses mainly on stateless, static network configurations, whereas we focus on stateful, dynamic applications.
+Our work is most closely related to Maple~\citep{Maple}.
+Like Maple, we implement policies in a general purpose language, but rather than reactively memoizing decisions, we generate forwarding decision proactively through symbolic analysis.
 
-%???
-
-\section*{Future Work and Conclusion}
+\section*{Conclusion}
 In this paper, we have introduced Oak, a SDN library for proactively switch forwarding rules

### 2015-05-11 draft.tex
pseudocode

@@
 
-\begin{figure*}[ht]
+\begin{figure}[ht]
 \begin{lstlisting}[mathescape]
-let compile (f: pkt $\rightarrow$ forwarding_decision) : policy =
-      push most general packet to (empty) stack
-      set the current location to be the root of an empty decision tree
+let compile (f: pkt $\rightarrow$ decision) : policy =
+      stack $\gets$ empty.push(new packet)
+      root $\gets$ new unexplored node
       while (stack is not empty)
-         pop the top of the stack as pkt
-         run f of pkt resulting in a forwarding decision fd and a refined packet pkt
-         record the forwarding decision for the symbolic packet pkt in the decision tree
-         set the current location to be the root of the decision tree and repeat
-
+          loc $\gets$ root
+          pkt $\gets$ stack.pop
+          fd $\gets$ f(pkt) 
+          loc $\gets$ Leaf fd
+      return root    
 \end{lstlisting}
@@
 \label{fig:compile-pseudo}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[ht]
-\begin{lstlisting}
+\begin{figure}[ht]
+\begin{lstlisting}[mathescape]
 let run (pol: policy) (inputs: examples) : unit =
-      build the symbolic rules for policy pol
-      generate the concrete rules with all relations empty
-      for each packets in input,
-          find the first rule matching the packet, and record the forwarding decision fd
-          if packet is sent to the controller then
-              evaluate the policy on the packet, potentially updating relations
-              regenerate concrete rules
-              forward the packet according the evaluated forwarding decision
-          else
-              forward the packet according to fd 
+      sr $\gets$ build_symbolic_rules(pol)
+      rel $\gets$ empty
+      cr $\gets$ build_concrete_rules(sr, rel)
+      for each packets pkt in input,
+      match evaluate_rules(cr, pkt) with
+       | Some fd $\to$ Forward(pkt, fd)
+       | None $\to$
+           (fd, rel) $\gets$ evaluate_policy(pkt)
+           cr $\gets$ build_concrete_rules(sr, rel)
+           Forward(pkt, fd)     
 \end{lstlisting}
@@
 \label{fig:run-pseudo}
-  \end{figure*}
+  \end{figure}
 

### 2015-05-11 draft.tex
please pull often to avoid overwriting changes

@@
 
-We introduce a high-level SDN library that can be embedded into existing general-purpose languages.
+We developed a high-level SDN library that can be embedded into existing general-purpose languages.
 Our library exposes an API that allows programmers to directly express behavioral intent, 
@@
 
-Ultimately, the performance requirements of networks dictate that the centralized controller(s) can not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency compared to those being handled by switches.
+Ultimately, the performance requirements of networks dictate that the centralized controller(s) should not handle most traffic since redirecting packets to be handled by the controller incurs orders of magnitude increased latency compared to those being handled by switches.
 For this reason, compiler technologies for network programming languages must be able to push much of the control plane functionality into the data plane switches themselves.
 
-For Maple, the authors show how to optimize the controller so that packets are handled without having to run the function on them.
+In Maple\cite{Maple}, the authors show how to optimize the controller to reduce the number of packets handled by running the function on them.
 Maple does so by providing a packet API capturing each access to packet fields, and memoizing the results of the function for subsequent packets sharing the accessed fields' values.
@@
 
-In this work, we develop a language-independent API and compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
+In this work, we develop a language-independent API together with a compilation strategy that, like Maple, allows programmers to write network programs with the full generality of a high-level, general-purpose programming language.
 However, we address each of the shortcomings in Maple listed above. Our contributions are the following:
@@
 \begin{enumerate}
-\item We describe a static analysis on the user-defined forwarding policy that pushes symbolic packets through the function to explore all possible paths of execution.
-This analysis is then used to proactively generate forwarding rules to install in the network.
-\item We provide a packet API that lets programmers ask if a field is in a particular range (e.g., IP prefix) rather than if the field equals a particular value.
+\item We describe a static analysis on the user-defined forwarding policy that explore all possible paths of execution by evaluation the function on symbolic packets.
+This analysis is then used to proactively generate a network controller and initial forwarding rules to install on the network switches.
+\item We provide a packet API that lets programmers ask if a field is in a particular range (e.g.
+an IP prefix) rather than if the field equals a particular value.
 This more expressive test works together with our analysis to better capture programmer intent when setting up forwarding decisions in the data plane.
@@
 
-The rest of the API, included in Figure~\ref{fig:build_api}, consists of functions used to interact with Red Oak.
+The rest of the API, included in Figure~\ref{fig:build_api}, consists of functions used to interact with \textit{Red Oak}.
 The \lstinline|compile| function is used to symbolically analyze a user-function for forwarding packets and generate, from this function, a decision tree that is captured as an abstract type called policy.
 The \lstinline|run| function simulates the compiled \lstinline|policy| on a series of example packets entering the network.
-This incorporates the Red Oak run time system to update forwarding decisions based on the global state.
+This incorporates the \textit{Red Oak} run time system to update forwarding decisions based on the global state.
 In the remainder of this section, we describe these functions in more detail.
@@
       match evaluate_rules(cr, pkt) with
-       | Some fd $\to$ Forward(pkt, fd)
+       | Some fd $\to$ forward(pkt, fd)
        | None $\to$
@@
            cr $\gets$ build_concrete_rules(sr, rel)
-           Forward(pkt, fd)     
+           forward(pkt, fd)     
 \end{lstlisting}
@@
   \end{figure}
+
+While we currently start the simulation with empty relation, \lstinline|run| could easily be extended to initialize the controller with pre-populated relations.
 
@@
 The Frenetic~\citep{Frenetic} family of languages provide domain-specific operators for composing packet processing functions together and compiling them to switches.
-Here we focus on compiling arbitrary programmer functions in a general-purpose language.
+Here, we focus on compiling arbitrary programmer functions in a general-purpose language.
 Merlin~\cite{Merlin} is a language for configuring networks based on path and quality of service constraints.
@@
 \section*{Conclusion}
-In this paper, we have introduced Oak, a SDN library for proactively switch forwarding rules
+In this paper, we have introduced \textit{Red Oak}, a SDN library for proactively switch forwarding rules
 from an arbitrary programmer-defined policy in a general purpose programming language.
-Oak provides programmers the abstraction that their centralized policy will run on each 
+\textit{Red Oak} provides programmers the abstraction that their centralized policy will run on each 
 new packet entering the network, while alleviating for the programmer need to consider
@@
 
-The Oak API exposes simple tests on packet fields as well as a global, mutable table data
+The \textit{Red Oak} API exposes simple tests on packet fields as well as a global, mutable table data
 structure for recording information about previous packets observed in the network.
@@
 packets will be handled, determining which packets must be observed by the controller and which 
-packets need not be. Finally, we describe our Oak prototype's run-time system that 
+packets need not be. Finally, we describe our \textit{Red Oak} prototype's run-time system that 
 specializes decisions after making concrete observations in the network that dictate forwarding policy.

### 2015-05-12 draft.tex
fixed ipsrc ipdst to Src and Dst

@@
 
-\lstset{morekeywords={in_range, in_rel, add, remove, Drop, Deliver, ForwardTo, Multicast, IpSrc, IpDst}}
+\lstset{morekeywords={in_range, in_rel, add, remove, Drop, Deliver, ForwardTo, Multicast, Src, Dst}}
 
@@
 f(pkt):
-  if in_range(pkt, src, 0, 10) then
-     if in_range pkt dst  [(0,10)] then
+  if in_range(pkt, Src, 0, 10) then
+     if in_range pkt Dst  [(0,10)] then
         Deliver
      else 
-        add pkt (IpDst, IpSrc) ``conn'';
+        add pkt (Dst, Src) ``conn'';
         Deliver
-  else if in_rel pkt (IpSrc,IpDst) ``conn'' then
+  else if in_rel pkt (Src,Dst) ``conn'' then
          Deliver
@@
   \begin{lstlisting}[mathescape]
-src $\in$ [0, 100] $\mapsto$ Controller
-src $\in$ [101, 1000], (src,dst) $\in$ conn  $\mapsto$ Deliver
-src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
+Src $\in$ [0, 100] $\mapsto$ Controller
+Src $\in$ [101, 1000], (Src,Dst) $\in$ conn  $\mapsto$ Deliver
+Src $\in$ [101, 1000], (Src,Dst) $\notin$ conn $\mapsto$ Drop
   \end{lstlisting}
@@
 \begin{lstlisting}[mathescape]
-src $\in$ [0, 100] $\mapsto$ Controller  
-src $\in$ [101, 1000] $\mapsto$ Drop  
+Src $\in$ [0, 100] $\mapsto$ Controller  
+Src $\in$ [101, 1000] $\mapsto$ Drop  
 \end{lstlisting}
@@
 
-  For example, a packing coming from \lstinline|src| 150 and \lstinline|dst| 10 would be matched by the second rule, and dropped on the switch.
-Subsequently, on a packet with \lstinline|src| 10 and \lstinline|dst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would be updated as follows:
+  For example, a packing coming from \lstinline|Src| 150 and \lstinline|Dst| 10 would be matched by the second rule, and dropped on the switch.
+Subsequently, on a packet with \lstinline|Src| 10 and \lstinline|Dst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would be updated as follows:
 
 \begin{lstlisting}[mathescape]
-src $\in$ [0, 100] $\mapsto$ Controller  
-src = 150, dst = 10 $\mapsto$ Deliver  
-src $\in$ [101, 1000] $\mapsto$ Drop  
+Src $\in$ [0, 100] $\mapsto$ Controller  
+Src = 150, Dst = 10 $\mapsto$ Deliver  
+Src $\in$ [101, 1000] $\mapsto$ Drop  
 \end{lstlisting}
@@
    \subsection*{Forwarding Decision}
-   % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to dst)
+   % can give constant decision (e.g. drop() or fwd(loc) to a certain location loc), or decision based on values of the packet (e.g. deliver() which forwards to Dst)
 	The programming model for \textit{Red Oak} is that each packet entering the network is mapped to a forwarding decision that describes where the packet should end up leaving the network.
@@
 
-let filter_src_range f r pkt =
- if in_range pkt IpSrc r then
+let filter_Src_range f r pkt =
+ if in_range pkt Src r then
     f pkt
@@
 
-   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_src_range| a partial policy, which, given a policy \lstinline|f| and a range \lstinline|r|, will forward all packets whose source IP is within range \lstinline|r| and drop the packet otherwise.
+   For example, we include in Figure~\ref{fig:ex-hole} the source code for \lstinline|filter_Src_range| a partial policy, which, given a policy \lstinline|f| and a range \lstinline|r|, will forward all packets whose source IP is within range \lstinline|r| and drop the packet otherwise.
 The ability to compose two or more policies in sequence like this arises many times in practice.
@@
 type pkt
-type field = IpSrc | IpDst 
+type field = Src | Dst 
 type fields = field list
@@
   \begin{lstlisting}[mathescape]
-src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
-src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
+Src $\in$ [101, 1000], (Src, Dst) $\in$ conn $\mapsto$ Deliver
+Src $\in$ [101, 1000], (Src, Dst) $\notin$ conn $\mapsto$ Drop
   \end{lstlisting}
@@
   Assume that the trusted connection relation ``conn'' contains the pairs 
-  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external src 150 can talk to internal dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
+  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external Src 150 can talk to internal Dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
 \begin{lstlisting}[mathescape]
-src $\in$ ([101, 1000] $\cap$ 150), dst = 10 $\mapsto$ Deliver
-src $\in$ ([101, 1000] $\cap$ 200), dst = 9 $\mapsto$ Deliver
-src $\in$ ([101, 1000] $\cap$ 8), dst = 150 $\mapsto$ Deliver
+Src $\in$ ([101, 1000] $\cap$ 150), Dst = 10 $\mapsto$ Deliver
+Src $\in$ ([101, 1000] $\cap$ 200), Dst = 9 $\mapsto$ Deliver
+Src $\in$ ([101, 1000] $\cap$ 8), Dst = 150 $\mapsto$ Deliver
  \end{lstlisting}
@@
 \begin{lstlisting}[mathescape]
-src = 150, dst = 10 $\mapsto$ Deliver
-src = 200, dst = 9 $\mapsto$ Deliver
+Src = 150, Dst = 10 $\mapsto$ Deliver
+Src = 200, Dst = 9 $\mapsto$ Deliver
  \end{lstlisting}
-Note the omission of the third concrete rule, resulting from \lstinline|ipsrc| having an empty range.
+Note the omission of the third concrete rule, resulting from \lstinline|Src| having an empty range.
 We can now concretize the second symbolic rule without doing any work, since checking for relation membership is already done by all the higher priority rules.
@@
 \begin{lstlisting}[mathescape]
-src = 150, dst = 10 $\mapsto$ Deliver
-src = 200, dst = 9 $\mapsto$ Deliver
-src $\in$ [101, 1000] $\mapsto$ Drop
+Src = 150, Dst = 10 $\mapsto$ Deliver
+Src = 200, Dst = 9 $\mapsto$ Deliver
+Src $\in$ [101, 1000] $\mapsto$ Drop
 \end{lstlisting}
@@
 However, these packets do not always modify the state of the controller.
-For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|ipsrc| of 10 and \lstinline|ipdst| of 150, any subsequent packet with the same \lstinline|ipsrc| and \lstinline|ipdst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent.
+For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|Src| of 10 and \lstinline|Dst| of 150, any subsequent packet with the same \lstinline|Src| and \lstinline|Dst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent.
 By using both local and global transformations on the decision tree, we can get closer to an optimal number of packet forwarded to the controller, where only packets having a real, meaningful effect are seen by the controller.
@@
 \begin{lstlisting}[mathescape]
-src $\in$ [0, 100], (dst, src) in conn $\mapsto$ Deliver
-src $\in$ [0, 100] $\mapsto$ Controller
-src $\in$ [101, 1000], (src,dst) $\in$ conn $\mapsto$ Deliver
-src $\in$ [101, 1000], (src,dst) $\notin$ conn $\mapsto$ Drop
+Src $\in$ [0, 100], (Dst, Src) in conn $\mapsto$ Deliver
+Src $\in$ [0, 100] $\mapsto$ Controller
+Src $\in$ [101, 1000], (Src,Dst) $\in$ conn $\mapsto$ Deliver
+Src $\in$ [101, 1000], (Src,Dst) $\notin$ conn $\mapsto$ Drop
 \end{lstlisting}
@@
 With empty relation, the concrete rules would be the same as with the unoptimized decision tree.
-However, after seeing, for example, packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the rules would be
+However, after seeing, for example, packet with \lstinline|Src| 10 and \lstinline|Dst| 150, the rules would be
 \begin{lstlisting}[mathescape]
-src = 10, dst = 150 $\mapsto$ Deliver  
-src $\in$ [0, 100] $\mapsto$ Controller  
-src = 150, dst = 10 $\mapsto$ Deliver  
-src $\in$ [101, 1000] $\mapsto$ Drop  
+Src = 10, Dst = 150 $\mapsto$ Deliver  
+Src $\in$ [0, 100] $\mapsto$ Controller  
+Src = 150, Dst = 10 $\mapsto$ Deliver  
+Src $\in$ [101, 1000] $\mapsto$ Drop  
 \end{lstlisting}
@@
 For example, we could remove \lstinline|add| node which are followed by \lstinline|remove| node for the same relation and tuple on all path without the change being observed by an \lstinline|in_rel| node.
-Similarly, we could detect and warn the user if an added tuple may never be observed by an \lstinline|in_rel| node, for example by switching \lstinline|ipsrc| and \lstinline|ipdst| in the add tuple of the stateful firewall example (see Figure~\ref{fig:firewallcode}).
+Similarly, we could detect and warn the user if an added tuple may never be observed by an \lstinline|in_rel| node, for example by switching \lstinline|Src| and \lstinline|Dst| in the add tuple of the stateful firewall example (see Figure~\ref{fig:firewallcode}).
 We leave the implementation of these optimizations as future work.

### 2015-05-12 draft.tex
..

@@
-\documentclass[preprint]{sigplanconf}
+\documentclass[nocopyrightspace]{sigplanconf}
 \usepackage[latin1]{inputenc}
@@
 
-\section*{Algorithic Policies Revisited}
+\section*{Algorithmic Policies Revisited}
 % Explains what algorithmic policies are, how maple was handling them and the difference with our model 
@@
    We provide several types of forwarding decisions to the programmer.
-The programmer can deliver a packet, according to the packet's desination, drop a packet, and forwarding a packet to a static location.
+The programmer can deliver a packet, according to the packet's destination, drop a packet, and forwarding a packet to a static location.
 Our API also allows for dynamic forwarding based on component of relation, for example to the last seen location of a host according to a MAC-learning controller (not shown in Fig~\ref{fig:range_api}).
@@
    % cross, ``holes''
-   As algorithmic policies under our model are functions of the host language, we inherit the latter's function space ability, allowing, amongst other things, for compositon of algorithmic policies and higher-order algorithmic policies.
+   As algorithmic policies under our model are functions of the host language, we inherit the latter's function space ability, allowing, among other things, for composition of algorithmic policies and higher-order algorithmic policies.
 
@@
 let compile (f: pkt $\rightarrow$ decision) : policy =
-      stack $\gets$ empty.push(new packet)
+      stack $\gets$ empty.push(new sym packet)
       root $\gets$ new unexplored node
@@
 To do so, we repeatedly run the algorithmic policy \lstinline|f| on symbol packets generated to explore every reachable path.
-We begin by calling \lstinline|f| on a symbolic packet covering the full range of value for each of its fields.
+We begin by calling \lstinline|f| on a symbolic packet covering the full range of values for each of its fields.
 
@@
 \begin{enumerate}
-  \item Create a new \lstinline|in_rel| branching node in the decision tree wit hsubtrees corresponding to \lstinline|true| and \lstinline|false|
+  \item Create a new \lstinline|in_rel| branching node in the decision tree with subtrees corresponding to \lstinline|true| and \lstinline|false|
   \item Tag the current packet with the symbolic constraint of being in the relation and return true
@@
 
-When \lstinline|compile| terminates, a complete decision tree has been traced for the input algoritmic policy.
+When \lstinline|compile| terminates, a complete decision tree has been traced for the input algorithmic policy.
 The next step is generate symbolic rules, mapping a symbolic packet to a forwarding decision.
@@
 For each tuple in the relation on which the symbolic rules depend, we create a new rule, refining the ranges of the relevant fields of the symbolic packet with the values of the tuple.
-We repeat this operation on the created list of symbolic rules until all dependant relations are taken care of, at which point the rules are considered concrete.
+We repeat this operation on the created list of symbolic rules until all dependent relations are taken care of, at which point the rules are considered concrete.
 Due to the prioritization of the resulting rules, we do not need to modify symbolic rules marked as not in a certain relation, as all the packets in the relation will have been matched by the positive rule before reaching this one.
@@
   Assume that the trusted connection relation ``conn'' contains the pairs 
-  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external Src 150 can talk to internal Dst 10 and so on. We concretize the above symbolic rules by expanding the first rule to:
+  \lstinline| {(150,10), (200, 9), (8, 150)}|, meaning external Src 150 can talk to internal Dst 10 and so on. We make the above symbolic rules concrete by expanding the first rule to:
 \begin{lstlisting}[mathescape]
@@
 Note the omission of the third concrete rule, resulting from \lstinline|Src| having an empty range.
-We can now concretize the second symbolic rule without doing any work, since checking for relation membership is already done by all the higher priority rules.
+We can now make the second symbolic rule concrete without doing any work, since checking for relation membership is already done by all the higher priority rules.
 The final concrete forwarding rules are now:
@@
        | None $\to$
-           (fd, rel) $\gets$ evaluate_policy(pkt)
+         (fd, rel) $\gets$ evaluate_policy(pol, pkt)
          cr $\gets$ build_concrete_rules(sr, rel)
@@
 We implemented in \textit{Red Oak} a local transformation adding an \lstinline|in_rel| node before \lstinline|add| or \lstinline|remove| nodes.
-In the case of \lstinline|add|, the \lstinline|add| node is omited in the true branch of \lstinline|in_rel|, such that packets whose fields are already in the relation will not be seen by the controller due to this \lstinline|add| node.
+In the case of \lstinline|add|, the \lstinline|add| node is omitted in the true branch of \lstinline|in_rel|, such that packets whose fields are already in the relation will not be seen by the controller due to this \lstinline|add| node.
 Similarly, we omit the \lstinline|remove| node in the false branch of \lstinline|in_rel|.
